---
title: "Analysis for tradition values"
output: 
  html_document: 
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

Two packages are not in CRAN and need different installation
* devtools::install_github("vjilmari/vjihelpers")
* install.packages("ggflags", repos = c("https://jimjam-slam.r-universe.dev","https://cloud.r-project.org")) 

```{r message=FALSE}

library(multid)
library(rio)
library(dplyr)
library(lme4)
library(emmeans)
library(vjihelpers)
library(MetBrewer)
library(ggplot2)
library(finalfit)
library(ggflags) 
library(lmerTest)
library(metafor)
library(ggpubr)
library(Hmisc)
library(r2mlm)
library(tidyr)
library(stringr)
library(apaTables)
library(tibble)


```

# Custom functions

```{r}
# toster function for equivalence testing

tost_z <- function(est, se, low, high, alpha = 0.10) {
  # z statistics
  z_low  <- (est - low)  / se     # H0: theta <= low  vs H1: theta > low
  z_high <- (est - high) / se     # H0: theta >= high vs H1: theta < high
  
  # one-sided p-values
  p_low  <- 1 - pnorm(z_low)
  p_high <- pnorm(z_high)
  
  # CI corresponding to TOST (1 - 2*alpha)
  z_crit <- qnorm(1 - alpha)
  ci_low  <- est - z_crit * se
  ci_high <- est + z_crit * se
  
  # equivalence decision
  equivalent <- (p_low < alpha) && (p_high < alpha)
  
  list(
    estimate     = est,
    se           = se,
    low_bound    = low,
    high_bound   = high,
    alpha        = alpha,
    z_low        = z_low,
    p_low        = p_low,
    z_high       = z_high,
    p_high       = p_high,
    ci_level     = 1 - 2*alpha,
    ci_lower     = ci_low,
    ci_upper     = ci_high,
    equivalent   = equivalent
  )
}


# TOST function for equivalence testing with t distribution

tost_t <- function(est, se, low, high, df, alpha = 0.10) {
  # t statistics
  t_low  <- (est - low)  / se   # H0: theta <= low  vs H1: theta > low
  t_high <- (est - high) / se   # H0: theta >= high vs H1: theta < high
  
  # one-sided p-values
  p_low  <- 1 - pt(t_low,  df = df)
  p_high <-     pt(t_high, df = df)
  
  # CI corresponding to TOST (1 - 2*alpha)
  t_crit <- qt(1 - alpha, df = df)
  ci_low  <- est - t_crit * se
  ci_high <- est + t_crit * se
  
  # equivalence decision
  equivalent <- (p_low < alpha) && (p_high < alpha)
  
  list(
    estimate   = est,
    se         = se,
    df         = df,
    low_bound  = low,
    high_bound = high,
    alpha      = alpha,
    t_low      = t_low,
    p_low      = p_low,
    t_high     = t_high,
    p_high     = p_high,
    ci_level   = 1 - 2 * alpha,
    ci_lower   = ci_low,
    ci_upper   = ci_high,
    equivalent = equivalent
  )
}

```


# Data

* Loads the preprocessed European Social Survey data file made within "Preparations_ESS" code

```{r}
load("../data/fdat.rdata")

```

* Include ISO data for country-names

```{r}
# read country names
ISO<-read.csv2("../data/ISO.csv")
```


# Exclusions

* exclude participants with missing values or gender

```{r}
value.vars<-
  c("con","tra",
  "ben","uni",
  "sdi","sti",
  "hed","ach",
  "pow","sec")

fdat$miss_values<-
  rowSums(is.na(fdat[,value.vars]))
table(fdat$miss_values)

fdat<-fdat %>%
  filter(miss_values==0 & !is.na(gndr.bin))
```


# Data for the analysis

* use the testing partition of the data set (diff_dat)
* basic descriptives

```{r}

diff_dat<-
  fdat

#diff_dat$age.c<-as.numeric(diff_dat$age.c)
#diff_dat$rlgdgr.c<-as.numeric(diff_dat$rlgdgr.c)
#diff_dat$eduyrs.c<-as.numeric(diff_dat$eduyrs.c)

# data exclusions (include countries with at least two rounds of data)
n_rounds<-diff_dat %>% group_by(cntry) %>%
  summarise(n_unique_essround = n_distinct(essround))

# join number of rounds to data
diff_dat<-
  left_join(x=diff_dat,
            y=n_rounds,
            by="cntry")

# filter only countries with more than 1 round and non-missing gender variable
diff_dat<-diff_dat %>%
  dplyr::filter(n_unique_essround>1 &
                  !is.na(gndr))

# cross-tabulate sample sizes and ESS rounds by country
table(diff_dat$essround,diff_dat$cntry)
# range of sample sizes
range(table(diff_dat$cntry))


# scale/standardize with SD pooled across all country-time-gender folds
cntry.tra<-diff_dat %>% group_by(cntry,essround) %>%
  summarise(tra.ctm=mean(tra,na.rm=T),
            tra.ctsd=sd(tra,na.rm=T)) %>%
  group_by(cntry) %>%
  summarise(tra.cm=mean(tra.ctm),
            tra.csd=mean(tra.ctsd)) 

grand_mean_tra<-mean(cntry.tra$tra.cm)
grand_sd_tra<-mean(cntry.tra$tra.csd)

# standardized
diff_dat$tra.z<-(diff_dat$tra-grand_mean_tra)/grand_sd_tra
hist(diff_dat$tra.z)

# recode time to start from 2002=0

diff_dat$year<-
  case_when(
    diff_dat$essround==1~2002,  
    diff_dat$essround==2~2004,
    diff_dat$essround==3~2006,
    diff_dat$essround==4~2008,
    diff_dat$essround==5~2010,
    diff_dat$essround==6~2012,
    diff_dat$essround==7~2014,
    diff_dat$essround==8~2016,
    diff_dat$essround==9~2018,
    diff_dat$essround==10~2020,
    diff_dat$essround==11~2023
  )

diff_dat$year.c<-diff_dat$year-2002


```

## Add GEI-variable

* GEI = Reverse-coded Gender Inequality Index (GII)


```{r}
GII<-read.csv("../data/GII.csv")

# Obtain GII only for ESS countries
GII_in_ESS_d<-
  GII %>%
  filter(ISO2 %in% unique(diff_dat$cntry))

# Make long format data of GII, so that country x year has own rows

long_GII_in_ESS_d <- GII_in_ESS_d %>%
  pivot_longer(
    cols = starts_with("gii_") & !matches("avg"),  # <-- exclude the "_avg" columns
    names_to = "year",
    values_to = "gii",
    names_prefix = "gii_"
  ) %>%
  mutate(year = as.integer(year)) %>%
  filter(year >= 2002 & year <= 2023) %>%
  mutate(gei = 1 - gii) %>%
  dplyr::select(ISO2, iso3, country, year, gii, gei, gii_2002_2023_avg, gei_2002_2023_avg)


# describe gei average
psych::describe(GII_in_ESS_d$gei_2002_2023_avg)
# one is missing, see which one
GII_in_ESS_d[is.na(GII_in_ESS_d$gei_2002_2023_avg),"country"]
# get means and SDs for standardizing
gei_mean<-mean(GII_in_ESS_d$gei_2002_2023_avg,na.rm=T)
gei_sd<-sd(GII_in_ESS_d$gei_2002_2023_avg,na.rm=T)

# standardize 
# year specific scores
long_GII_in_ESS_d$gei.z<-(long_GII_in_ESS_d$gei-gei_mean)/gei_sd

# get the country means to a separate frame

GII_country_means<-
  long_GII_in_ESS_d %>% group_by(ISO2) %>%
  summarise(gei.cm=mean(gei,na.rm=T),
            gei.z.cm=mean(gei.z,na.rm=T))


long_GII_in_ESS_d<-left_join(
  x=long_GII_in_ESS_d,
  y=GII_country_means,
  by="ISO2"
)


# center both the raw score and standardized scores
long_GII_in_ESS_d$gei.cmc<-(long_GII_in_ESS_d$gei-long_GII_in_ESS_d$gei.cm)
long_GII_in_ESS_d$gei.z.cmc<-(long_GII_in_ESS_d$gei.z-long_GII_in_ESS_d$gei.z.cm)

#View(long_GII_in_ESS_d)

# averages across 2002-2023
#long_GII_in_ESS_d$gei_2002_2023_avg.z<-(long_GII_in_ESS_d$gei_2002_2023_avg-gei_mean)/gei_sd

#long_GII_in_ESS_d$gei.cmc<-long_GII_in_ESS_d$gei-long_GII_in_ESS_d$gei_2002_2023_avg

# add year to ESS data-frame


# link datasets by country and year

diff_dat<-
  left_join(
    x=diff_dat,
    y=long_GII_in_ESS_d[,c("ISO2","year","gei","gei.cm","gei.cmc",
                           "gei.z","gei.z.cm","gei.z.cmc")],
    by=c("cntry"="ISO2","year")
)

```


## Add GDI-variable

* GDI = Gender Development Index

```{r}
GDI<-read.csv("../data/GDI.csv")

# Obtain GDI only for ESS countries
GDI_in_ESS_d<-
  GDI %>%
  filter(ISO2 %in% unique(diff_dat$cntry))

# Make long format data of GDI, so that country x year has own rows

long_GDI_in_ESS_d <- GDI_in_ESS_d %>%
  pivot_longer(
    cols = starts_with("gdi_") & !matches("avg"),  # <-- exclude the "_avg" columns
    names_to = "year",
    values_to = "gdi",
    names_prefix = "gdi_"
  ) %>%
  mutate(year = as.integer(year)) %>%
  filter(year >= 2002 & year <= 2023) %>%
  dplyr::select(ISO2, iso3, country, year, gdi, gdi_2002_2023_avg)

# describe gdi average
psych::describe(GDI_in_ESS_d$gdi_2002_2023_avg)
# get means and SDs for standardizing
gdi_mean<-mean(GDI_in_ESS_d$gdi_2002_2023_avg,na.rm=T)
gdi_sd<-sd(GDI_in_ESS_d$gdi_2002_2023_avg,na.rm=T)

# standardize 
# year specific scores
long_GDI_in_ESS_d$gdi.z<-(long_GDI_in_ESS_d$gdi-gdi_mean)/gdi_sd

# get the country means to a separate frame

GDI_country_means<-
  long_GDI_in_ESS_d %>% group_by(ISO2) %>%
  summarise(gdi.cm=mean(gdi,na.rm=T),
            gdi.z.cm=mean(gdi.z,na.rm=T))

long_GDI_in_ESS_d<-left_join(
  x=long_GDI_in_ESS_d,
  y=GDI_country_means,
  by="ISO2"
)


# center both the raw score and standardized scores
long_GDI_in_ESS_d$gdi.cmc<-(long_GDI_in_ESS_d$gdi-long_GDI_in_ESS_d$gdi.cm)
long_GDI_in_ESS_d$gdi.z.cmc<-(long_GDI_in_ESS_d$gdi.z-long_GDI_in_ESS_d$gdi.z.cm)

# link datasets by country and year

diff_dat<-
  left_join(
    x=diff_dat,
    y=long_GDI_in_ESS_d[,c("ISO2","year","gdi","gdi.cm","gdi.cmc",
                           "gdi.z","gdi.z.cm","gdi.z.cmc")],
    by=c("cntry"="ISO2","year")
  )

```

## Add GDP-variable

* log_GDP = Logarithm of GDP per capita, PPP (constant 2017 international $)


```{r}
GDP<-read.csv("../data/GDP_processed.csv")

# Obtain GDP only for ESS countries
GDP_in_ESS_d<-
  GDP %>%
  filter(ISO2 %in% unique(diff_dat$cntry))

# First, rename those X2000..YR2000. style columns to simple "gdp_2000" etc.

GDP_in_ESS_d <- GDP_in_ESS_d %>%
  rename_with(
    ~ str_replace_all(., "^X(\\d{4})..YR\\1\\.$", "gdp_\\1"),
    starts_with("X")
  )

# GDP_in_ESS_d
# Make long format data of GDP, so that country x year has own rows

long_GDP_in_ESS_d <- GDP_in_ESS_d %>%
  pivot_longer(
    cols = starts_with("gdp_") & !matches("avg"),  # <-- exclude the "_avg" columns
    names_to = "year",
    values_to = "gdp",
    names_prefix = "gdp_"
  ) %>%
  mutate(year = as.integer(year)) %>%
  filter(year >= 2002 & year <= 2023) %>%
  dplyr::select(ISO2, Country.Name, year, gdp, gdp_2002_2023_avg,log_gdp_2002_2023_avg) %>%
  mutate(log_gdp=log(gdp))

#View(long_GDP_in_ESS_d)

# describe gdp average
psych::describe(GDP_in_ESS_d$gdp_2002_2023_avg)
psych::describe(GDP_in_ESS_d$log_gdp_2002_2023_avg)

# get means and SDs for standardizing
log_gdp_mean<-mean(GDP_in_ESS_d$log_gdp_2002_2023_avg,na.rm=T)
log_gdp_sd<-sd(GDP_in_ESS_d$log_gdp_2002_2023_avg,na.rm=T)

# standardize 
# year specific scores
long_GDP_in_ESS_d$log_gdp.z<-
  (long_GDP_in_ESS_d$log_gdp-log_gdp_mean)/log_gdp_sd

# get the country means to a separate frame

GDP_country_means<-
  long_GDP_in_ESS_d %>% group_by(ISO2) %>%
  summarise(gdp.cm=mean(gdp,na.rm=T),
            #gdp.z.cm=mean(gdp.z,na.rm=T),
            log_gdp.cm=mean(log_gdp,na.rm=T),
            log_gdp.z.cm=mean(log_gdp.z,na.rm=T))

long_GDP_in_ESS_d<-left_join(
  x=long_GDP_in_ESS_d,
  y=GDP_country_means,
  by="ISO2"
)


# center both the raw score and standardized scores
long_GDP_in_ESS_d$log_gdp.cmc<-(long_GDP_in_ESS_d$log_gdp-long_GDP_in_ESS_d$log_gdp.cm)
long_GDP_in_ESS_d$log_gdp.z.cmc<-(long_GDP_in_ESS_d$log_gdp.z-long_GDP_in_ESS_d$log_gdp.z.cm)

# link datasets by country and year

diff_dat<-
  left_join(
    x=diff_dat,
    y=long_GDP_in_ESS_d[,c("ISO2","year","log_gdp","log_gdp.cm","log_gdp.cmc",
                           "log_gdp.z","log_gdp.z.cm","log_gdp.z.cmc")],
    by=c("cntry"="ISO2","year")
  )

  
```

## Add GGGI-variable

* GGGI = Global Gender Gap Index


```{r}
GGGI<-import("../data/GGGI.xlsx")

# check if all ESS countries are in the log_GGGI data
table(unique(diff_dat$cntry) %in% GGGI$ISO2)

# Obtain GGGI only for ESS countries
GGGI_in_ESS_d<-
  GGGI %>%
  filter(ISO2 %in% unique(diff_dat$cntry))

# Make long format data of GGGI, so that country x year has own rows

long_GGGI_in_ESS_d <- GGGI_in_ESS_d %>%
  pivot_longer(
    cols = starts_with("gggi_") & !matches("avg"),  # <-- exclude the "_avg" columns
    names_to = "year",
    values_to = "gggi",
    names_prefix = "gggi_"
  ) %>%
  mutate(year = as.integer(year)) %>%
  filter(year >= 2002 & year <= 2023) %>%
  dplyr::select(ISO2, cname, year, gggi, GGGI_2002_2023_avg)

# describe gggi average
psych::describe(GGGI_in_ESS_d$GGGI_2002_2023_avg)
# get means and SDs for standardizing
gggi_mean<-mean(GGGI_in_ESS_d$GGGI_2002_2023_avg,na.rm=T)
gggi_sd<-sd(GGGI_in_ESS_d$GGGI_2002_2023_avg,na.rm=T)

# standardize 
# year specific scores
long_GGGI_in_ESS_d$gggi.z<-(long_GGGI_in_ESS_d$gggi-gggi_mean)/gggi_sd

# get the country means to a separate frame

GGGI_country_means<-
  long_GGGI_in_ESS_d %>% group_by(ISO2) %>%
  summarise(gggi.cm=mean(gggi,na.rm=T),
            gggi.z.cm=mean(gggi.z,na.rm=T))

long_GGGI_in_ESS_d<-left_join(
  x=long_GGGI_in_ESS_d,
  y=GGGI_country_means,
  by="ISO2"
)

# center both the raw score and standardized scores
long_GGGI_in_ESS_d$gggi.cmc<-
  (long_GGGI_in_ESS_d$gggi-long_GGGI_in_ESS_d$gggi.cm)

long_GGGI_in_ESS_d$gggi.z.cmc<-
  (long_GGGI_in_ESS_d$gggi.z-long_GGGI_in_ESS_d$gggi.z.cm)

# link datasets by country and year

diff_dat<-
  left_join(
    x=diff_dat,
    y=long_GGGI_in_ESS_d[,c("ISO2","year","gggi","gggi.cm","gggi.cmc",
                           "gggi.z","gggi.z.cm","gggi.z.cmc")],
    by=c("cntry"="ISO2","year")
  )

```

## Country-year-gender dataframe

Frame with one row for country-year-gender

```{r}
diff_dat_cntry_year<-
  diff_dat %>% 
  group_by(cntry,year,gndr.c) %>%
  dplyr::summarise(n=n(),
                   n_wt=sum(pspwght),
                   tra.z.wt=weighted.mean(x=tra.z,w=pspwght),
                   tra.z=mean(tra.z),
                   gei.z=mean(gei.z,na.rm=T),
                   gei.z.cm=mean(gei.z.cm,na.rm=T),
                   gei.z.cmc=mean(gei.z.cmc,na.rm=T),
                   gdi.z=mean(gdi.z,na.rm=T),
                   gdi.z.cm=mean(gdi.z.cm,na.rm=T),
                   gdi.z.cmc=mean(gdi.z.cmc,na.rm=T),
                   gggi.z=mean(gggi.z,na.rm=T),
                   gggi.z.cm=mean(gggi.z.cm,na.rm=T),
                   gggi.z.cmc=mean(gggi.z.cmc,na.rm=T),
                   log_gdp.z=mean(log_gdp.z,na.rm=T),
                   log_gdp.z.cm=mean(log_gdp.z.cm,na.rm=T),
                   log_gdp.z.cmc=mean(log_gdp.z.cmc,na.rm=T),
                   gei=mean(gei,na.rm=T),
                   gei.cm=mean(gei.cm,na.rm=T),
                   gei.cmc=mean(gei.cmc,na.rm=T),
                   gdi=mean(gdi,na.rm=T),
                   gdi.cm=mean(gdi.cm,na.rm=T),
                   gdi.cmc=mean(gdi.cmc,na.rm=T),
                   gggi=mean(gggi,na.rm=T),
                   gggi.cm=mean(gggi.cm,na.rm=T),
                   gggi.cmc=mean(gggi.cmc,na.rm=T),
                   log_gdp=mean(log_gdp,na.rm=T),
                   log_gdp.cm=mean(log_gdp.cm,na.rm=T),
                   log_gdp.cmc=mean(log_gdp.cmc,na.rm=T))
```


# Descriptive statistics

## Country-specific descriptives

```{r}

# sample sizes from weights

cntry_n_frame<-
  diff_dat %>% group_by(cntry) %>%
  summarise('n ESS rounds' = mean(n_unique_essround),
            n=round(sum(pspwght),0))

# tradition

cntry_tra_frame<-
  diff_dat %>% group_by(cntry,essround) %>%
  summarise('tra M' = weighted.mean(x=tra.z,w=pspwght),
            'tra SD' = sqrt(wtd.var(tra.z,pspwght))) %>%
  group_by(cntry) %>%
  summarise('tra M' = mean(x=`tra M`),
            'tra SD'= mean(x=`tra SD`))

cntry_tra_women_frame<-
  diff_dat %>%
  filter(gndr.c==-0.5) %>%
  group_by(cntry,essround) %>%
  summarise('tra M' = weighted.mean(x=tra.z,w=pspwght),
            'tra SD' = sqrt(wtd.var(tra.z,pspwght))) %>%
  group_by(cntry) %>%
  summarise('tra M Women' = mean(x=`tra M`),
            'tra SD Women'= mean(x=`tra SD`))

cntry_tra_men_frame<-
  diff_dat %>%
  filter(gndr.c==0.5) %>%
  group_by(cntry,essround) %>%
  summarise('tra M' = weighted.mean(x=tra.z,w=pspwght),
            'tra SD' = sqrt(wtd.var(tra.z,pspwght))) %>%
  group_by(cntry) %>%
  summarise('tra M Men' = mean(x=`tra M`),
            'tra SD Men'= mean(x=`tra SD`))


# link n and tra datasets

desc_frame<-
  left_join(
    x=cntry_n_frame,
    y=cntry_tra_frame,
    by="cntry"
  )

desc_frame<-
  left_join(
    x=desc_frame,
    y=cntry_tra_women_frame,
    by="cntry"
  )

desc_frame<-
  left_join(
    x=desc_frame,
    y=cntry_tra_men_frame,
    by="cntry"
  )

# Add country-specific differences
desc_frame$D<-desc_frame$`tra M Men`-desc_frame$`tra M Women`

desc_frame
# add country-level measures

desc_frame<-
  left_join(
    x=desc_frame,
    y=GII_country_means[,c("ISO2","gei.cm")],
    by=c("cntry"="ISO2")
  )

desc_frame<-
  left_join(
    x=desc_frame,
    y=GGGI_country_means[,c("ISO2","gggi.cm")],
    by=c("cntry"="ISO2")
  )

desc_frame<-
  left_join(
    x=desc_frame,
    y=GDI_country_means[,c("ISO2","gdi.cm")],
    by=c("cntry"="ISO2")
  )


desc_frame<-
  left_join(
    x=desc_frame,
    y=GDP_country_means[,c("ISO2","gdp.cm")],
    by=c("cntry"="ISO2")
  )

desc_frame<-left_join(
  x=desc_frame,
  y=ISO[,c("ISO2","CLDR")],
  by=c("cntry"="ISO2")
)

# finalize the formatted table

cntry_desc_tbl<-
  desc_frame %>%
  dplyr::select(
    Country = CLDR,
    `n ESS rounds`,
    n,
    `tra M`, `tra SD`,
    `tra M Women`, `tra SD Women`,
    `tra M Men`, `tra SD Men`,
    D,
    GEI = gei.cm,
    GGGI = gggi.cm,
    GDI = gdi.cm,
    GDP = gdp.cm
  ) %>%
  mutate(
    across(
      .cols = -c(Country, `n ESS rounds`, n, GDP) & where(is.numeric),
      .fns  = ~ round_tidy(.x, 2)
    )
  ) %>%
  mutate(
    across(
      .cols = GDP,
      .fns  = ~ round_tidy(.x, 0)
    )
  )
print(cntry_desc_tbl,n=35)

export(cntry_desc_tbl,"../results/tra/cntry_desc_tbl.xlsx",overwrite=T)
```

## Country-level correlations table

```{r}
cor_frame<-
  desc_frame %>%
  dplyr::select(
    VBMT=`tra M`,
    VBMT_Women=`tra M Women`,
    VBMT_Men=`tra M Men`,
    D = D,
    GEI = gei.cm,
    GGGI = gggi.cm,
    GDI = gdi.cm,
    GDP = gdp.cm
  ) %>%
  mutate(
    log_GDP=log(GDP)
  ) %>%
  dplyr::select(-GDP)

apa.cor.table(
  data=cor_frame,
  filename = "../results/tra/CorTable1.doc",
  #table.number = NA,
  show.conf.interval = FALSE,
  show.sig.stars = F,
  landscape = F
)
```


# Analysis

Following preregistration: https://osf.io/7cags?view_only=f3e97a78271e46bfafb9e20ac8d35bb1 

## mod0: Random intercept model

```{r}

mod0<-lmer(tra.z~(1|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod0)
getVC(mod0)
r2mlm(mod0,bargraph = F)
```

## mod1: Gender fixed effect

```{r}

mod1<-lmer(tra.z~gndr.c+(1|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod1)
getFE(mod1,round=3)
getVC(mod1)
r2mlm(mod1,bargraph = F)
```

## mod2: Gender fixed and random effect

* Include random effect correlation by default

```{r}

mod2<-lmer(tra.z~gndr.c+(gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2)
getFE(mod2,round=3)
getVC(mod2)
r2mlm(mod2,bargraph = F)
anova(mod1,mod2)
lvl2_var_cond_lvl1(mod2,lvl1.var = "gndr.c",lvl1.values = c(0.5,-0.5))
```

* Test for random effect correlation

```{r}
mod2_norecov<-lmer(tra.z~gndr.c+(gndr.c||cntry),data=diff_dat,REML=F,weights = pspwght,
                   control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_norecov)
getFE(mod2_norecov,round=3)
getVC(mod2_norecov)

anova(mod2_norecov,mod2)
```


## mod2 with Gender-equality index (GEI)

```{r}
mod2_GEI<-lmer(tra.z~gndr.c+gei.z.cm+gndr.c:gei.z.cm+
                 (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_GEI)
getFE(mod2_GEI,round=3)
getVC(mod2_GEI)
r2mlm(mod2_GEI,bargraph = F)
```

### Deconstructed associations

```{r}
t1<-Sys.time()
ddsc_mod2_GEI<-
  ddsc_ml(model = mod2_GEI,
          predictor = "gei.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)
t2<-Sys.time()
t2-t1
round(ddsc_mod2_GEI$vpc_at_moderator_values,3)
round(ddsc_mod2_GEI$descriptives,3)
round(ddsc_mod2_GEI$results,3)
round(ddsc_mod2_GEI$re_cov_test,3)

# Structural path model
d_GEI<-ddsc_mod2_GEI$ddsc_sem_fit$data

ddsc_sem_GEI<-
  ddsc_sem(data=d_GEI,x = "gei.z.cm",y1="means_y2",
           y2="means_y1",q_sesoi = .10)
round(ddsc_sem_GEI$results,3)
round(ddsc_sem_GEI$variance_test,3)

## random intercept mlm with double-entries for each country (men and women)

d_GEI_long <- d_GEI %>%
  # move row names into a column
  rownames_to_column("cntry") %>%
  # pivot only the means_y1 / means_y2 columns
  pivot_longer(
    cols = c(means_y1, means_y2),
    names_to = "y",
    values_to = "means"
  ) %>%
  # create a sgender column based on y1/y2
  mutate(
    gndr.c = case_when(
      y == "means_y1" ~ -0.5,
      y == "means_y2" ~ 0.5
    )
  ) %>%
  dplyr::select(cntry,gei.z.cm,means,gndr.c)


ddsc_mod2_GEI_ri<-
  ddsc_ml(data=data.frame(d_GEI_long),predictor = "gei.z.cm",
          moderator = "gndr.c",
        DV = "means",lvl2_unit = "cntry",
        moderator_values = c(-0.5,0.5))
round(ddsc_mod2_GEI_ri$results,3)

# country-time multilevel model


mod2_GEI_cntry_year<-
  lmer(tra.z.wt~gndr.c+gei.z.cm+gndr.c:gei.z.cm+
                 (gndr.c|cntry),data=diff_dat_cntry_year,REML=F,
               control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_GEI_cntry_year)
getFE(mod2_GEI_cntry_year,round=3)
getVC(mod2_GEI_cntry_year)
r2mlm(mod2_GEI,bargraph = F)

ddsc_mod2_GEI_cntry_year<-
  ddsc_ml(model = mod2_GEI_cntry_year,
          predictor = "gei.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)

round(ddsc_mod2_GEI_cntry_year$vpc_at_moderator_values,3)
round(ddsc_mod2_GEI_cntry_year$descriptives,3)
round(ddsc_mod2_GEI_cntry_year$results,3)
round(ddsc_mod2_GEI_cntry_year$re_cov_test,3)

```

### Comparisons across models

```{r}
# full multilevel model (individuals as entries)
round(ddsc_mod2_GEI$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# country-level structural path model (country-gender means as entries)
round(ddsc_sem_GEI$results[c("r_xy1","r_xy2","b_11","b_21","q_b11_b21","diff_b11_b21"),],4)
# multilevel model (country-gender means as entries)
round(ddsc_mod2_GEI_ri$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# multilevel model (country-year-gender means as entries)
round(ddsc_mod2_GEI_cntry_year$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)

```


### Bootstrap and equivalence test

Takes a lot of time

```{r}

t1<-Sys.time()
mod2_GEI_booted_fixef <-
  lme4::bootMer(
    x = mod2_GEI,
    FUN = lme4::fixef,
    nsim = 1000,
    use.u = FALSE,
    seed = 12345,
    type = c("parametric"),
    verbose = FALSE
  )
t2<-Sys.time()
t2-t1
```


```{r}
# obtain all the bootstrap estimates
mod2_GEI_boot_est <- data.frame(mod2_GEI_booted_fixef$t)

# calculate estimates
mod2_GEI_boot_est$w11<-mod2_GEI_boot_est$gei.z.cm+(-0.5)*mod2_GEI_boot_est$gndr.c.gei.z.cm
mod2_GEI_boot_est$w21<-mod2_GEI_boot_est$gei.z.cm+(0.5)*mod2_GEI_boot_est$gndr.c.gei.z.cm
mod2_GEI_boot_est$b11<-mod2_GEI_boot_est$w11/ddsc_mod2_GEI$SDs["SD_pooled"]
mod2_GEI_boot_est$b21<-mod2_GEI_boot_est$w21/ddsc_mod2_GEI$SDs["SD_pooled"]
mod2_GEI_boot_est$r_xy1<-mod2_GEI_boot_est$w11/ddsc_mod2_GEI$SDs["SD_y1"]
mod2_GEI_boot_est$r_xy2<-mod2_GEI_boot_est$w21/ddsc_mod2_GEI$SDs["SD_y2"]
mod2_GEI_boot_est$q_b<-atanh(mod2_GEI_boot_est$b11)-atanh(mod2_GEI_boot_est$b21)
mod2_GEI_boot_est$q<-atanh(mod2_GEI_boot_est$r_xy1)-atanh(mod2_GEI_boot_est$r_xy2)

# Calculate bootstrap summary statistics
mod2_GEI_boot_results <- t(as.data.frame(sapply(
  mod2_GEI_boot_est,
  function(x) {
    c(
      Estimate = mean(x, na.rm = TRUE),
      SE = stats::sd(x, na.rm = TRUE),
      stats::quantile(x, c((1 - .95) / 2,
                           1 - (1 - .95) / 2), na.rm = TRUE)
    )
  }
)))

mod2_GEI_boot_results

# equivalence test for q_b
tost_z(est=mod2_GEI_boot_results["q_b","Estimate"],
       se=mod2_GEI_boot_results["q_b","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

# equivalence test for q
tost_z(est=mod2_GEI_boot_results["q","Estimate"],
       se=mod2_GEI_boot_results["q","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

```



### Figure 

```{r}

# plot the results

# start with obtaining predicted values for means and differences

# refit reduced and full models with GGGI in original scale


mod2_GEI_unstd<-lmer(tra.z~gndr.c+gei.cm+gndr.c:gei.cm+
                 (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
               control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))

mod2_GEI_unstd_red<-lmer(tra.z~gndr.c+
                       (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
                     control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))


p<-
  emmip(
    mod2_GEI_unstd, 
    gndr.c ~ gei.cm,
    at=list(gndr.c = c(-0.5,0.5),
            gei.cm=
              seq(from=round(range(GII_country_means$gei.cm,na.rm=T)[1],2),
                  to=round(range(GII_country_means$gei.cm,na.rm=T)[2],2),
                  by=0.001)),
    plotit=F,CIs=T,lmerTest.limit = 1e6,disable.pbkrtest=T)

p$gndr.c<-p$tvar
levels(p$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.comp<-min(p$LCL)
max.y.comp<-max(p$UCL)

# Men and Women mean distributions

p3<-coefficients(mod2_GEI_unstd_red)$cntry
p3<-cbind(rbind(p3,p3),weight=rep(c(-0.5,0.5),each=nrow(p3)))
p3$xvar<-p3$`(Intercept)`+p3$gndr.c*p3$weight
p3$gndr.c<-as.factor(p3$weight)
levels(p3$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.mean.distr<-min(p3$xvar)
max.y.mean.distr<-max(p3$xvar)


# obtain the coefs for the gndr.c-effect (difference) as function of gei.cm

p2<-data.frame(
  emtrends(mod2_GEI_unstd,var="gndr.c",
           specs="gei.cm",
           at=list(#gndr.c = c(-0.5,0.5),
             gei.cm=
               seq(from=round(range(GII_country_means$gei.cm,na.rm=T)[1],2),
                   to=round(range(GII_country_means$gei.cm,na.rm=T)[2],2),
                   by=0.001)),
           lmerTest.limit = 1e6,disable.pbkrtest=T))

p2$yvar<-p2$gndr.c.trend
p2$xvar<-p2$gei.cm
p2$LCL<-p2$lower.CL
p2$UCL<-p2$upper.CL

# obtain min and max for aligned plots
min.y.diff<-min(p2$LCL)
max.y.diff<-max(p2$UCL)

# difference score distribution

p4<-coefficients(mod2_GEI_unstd_red)$cntry
p4$xvar=(+1)*p4$gndr.c

# obtain mix and max for aligned plots

min.y.diff.distr<-min(p4$xvar)
max.y.diff.distr<-max(p4$xvar)

# define mins and maxs

min.y.pred<-
  ifelse(min.y.comp<min.y.mean.distr,min.y.comp,min.y.mean.distr)

max.y.pred<-
  ifelse(max.y.comp>max.y.mean.distr,max.y.comp,max.y.mean.distr)

min.y.narrow<-
  ifelse(min.y.diff<min.y.diff.distr,min.y.diff,min.y.diff.distr)

max.y.narrow<-
  ifelse(max.y.diff>max.y.diff.distr,max.y.diff,max.y.diff.distr)

# Figures 

# p1

# scaled simple effects to the plot
pvals<-round_tidy(ddsc_mod2_GEI$results[6:7,"p.value"],3)

ests<-
  round_tidy(ddsc_mod2_GEI$results[6:7,"estimate"],2)

coef1<-paste0("std. b11 = ",ests[1],", p = ",pvals[1])
coef2<-paste0("std. b21 = ",ests[2],", p = ",pvals[2])
coefs<-data.frame(gndr.c=c("Women","Men"),
                  coef=c(coef1,coef2))

coef_q<-paste0("Cohen's q = ",round_tidy(ddsc_mod2_GEI$results["q_b11_b21","estimate"],2),", p = ",
               substr(round_tidy(ddsc_mod2_GEI$results["interaction","p.value"],3),2,5))  
# prediction plot for difference score

pvals2<-round_tidy(ddsc_mod2_GEI$results["interaction","p.value"],3)

ests2<-
  round_tidy(-1*ddsc_mod2_GEI$results["r_xy1y2","estimate"],2)

coefs2<-paste0("difference score correlation = ",ests2,", p = ",pvals2)
# attempt to make a flag-plot

flag_points_GEI<-coefficients(mod2_GEI_unstd_red)$cntry

flag_points_GEI$women<-flag_points_GEI$`(Intercept)`+(-0.5)*flag_points_GEI$gndr.c
flag_points_GEI$men<-flag_points_GEI$`(Intercept)`+(0.5)*flag_points_GEI$gndr.c
flag_points_GEI$mean_level<-flag_points_GEI$`(Intercept)`
flag_points_GEI$difference<-flag_points_GEI$men-flag_points_GEI$women
flag_points_GEI$cntry<-rownames(flag_points_GEI)

flag_points_GEI<-
  left_join(x=flag_points_GEI,
            y=GII_country_means[,c("ISO2","gei.cm")],by=c("cntry"="ISO2"))
#flag_points_GEI

flag_points_GEI_long<-
  data.frame(mean_level=c(flag_points_GEI$women,
                          flag_points_GEI$men),
             gei.cm=c(flag_points_GEI$gei.cm,
                      flag_points_GEI$gei.cm),
             cntry=c(flag_points_GEI$cntry,
                     flag_points_GEI$cntry),
             gndr.c=rep(c("Women","Men"),each=nrow(flag_points_GEI)
             ))


p1.tra.flags<-
  ggplot(p,aes(y=yvar,x=gei.cm,color=gndr.c))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("Gender Equality Index (Average 2002-2023)")+
  #ylim(c(min.y.pred,max.y.pred))+
  #xlim(c(0.60,1.00))+
  ylab("Value tradition (Average 2002-2023)")+
  scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "top",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(linewidth = 0.5, linetype = 2,
                                          colour = "gray"))+
  geom_text(data = coefs,show.legend=F,
            aes(label=coef,x=0.65,
                y=c(-0.35,-0.40),size=14,hjust="left"))+
  geom_text(inherit.aes=F,aes(x=0.65,y=-0.45,
                              label=coef_q,size=14,hjust="left"),
            show.legend=F)+
  geom_point(data=flag_points_GEI_long,size=8,alpha=0.30,
             aes(x=gei.cm,y=mean_level))+
  geom_line(data=flag_points_GEI_long,aes(group = cntry,x=gei.cm,y=mean_level),
            color="black",linetype=2)+
  geom_flag(data=flag_points_GEI_long,show.legend=F,
            aes(country=tolower(cntry),size=14,x=gei.cm,y=mean_level))

p2.tra.flags<-ggplot(p2,aes(y=yvar,x=gei.cm))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("Gender Equality Index (Average 2002-2023)")+
  #xlim(c(0.60,1.00))+
  #ylim(c(min.y.narrow,max.y.narrow))+
  ylab("Difference in value tradition")+
  #scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "right",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(linewidth = 0.5, linetype = 2,
                                          colour = "gray"))+
  #geom_text(coef2,aes(x=0.63,y=min(p2$LCL)))
  geom_text(data = data.frame(coefs2),show.legend=F,
            aes(label=coefs2,x=0.70,
                y=c(round(min(p2$LCL),2))+0.05,size=18,hjust="left"))+
  geom_flag(data=flag_points_GEI,show.legend=F,
            aes(country=tolower(cntry),size=18,x=gei.cm,y=difference))

pflag_comb<-
  ggarrange(p1.tra.flags,p2.tra.flags,align = "v",
            ncol=1,nrow=2,heights=c(2,1))
pflag_comb


png(filename = 
      "../results/tra/GEI_flags.png",
    units = "cm",
    width = 21.0,height=29.7*(3/4),res = 300)
pflag_comb
dev.off()

```

## mod2 with Gender-equality index (GGGI)

```{r}
mod2_GGGI<-lmer(tra.z~gndr.c+gggi.z.cm+gndr.c:gggi.z.cm+
                 (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_GGGI)
getFE(mod2_GGGI,round=3)
getVC(mod2_GGGI)
r2mlm(mod2_GGGI,bargraph = F)
```


### Deconstructed associations

```{r}
t1<-Sys.time()
ddsc_mod2_GGGI<-
  ddsc_ml(model = mod2_GGGI,
          predictor = "gggi.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)
t2<-Sys.time()
t2-t1
round(ddsc_mod2_GGGI$vpc_at_moderator_values,3)
round(ddsc_mod2_GGGI$descriptives,3)
round(ddsc_mod2_GGGI$results,3)
round(ddsc_mod2_GGGI$re_cov_test,3)

# Structural path model
d_GGGI<-ddsc_mod2_GGGI$ddsc_sem_fit$data

ddsc_sem_GGGI<-
  ddsc_sem(data=d_GGGI,x = "gggi.z.cm",y1="means_y2",
           y2="means_y1",q_sesoi = .10)
round(ddsc_sem_GGGI$results,3)
round(ddsc_sem_GGGI$variance_test,3)

## random intercept mlm with double-entries for each country (men and women)

d_GGGI_long <- d_GGGI %>%
  # move row names into a column
  rownames_to_column("cntry") %>%
  # pivot only the means_y1 / means_y2 columns
  pivot_longer(
    cols = c(means_y1, means_y2),
    names_to = "y",
    values_to = "means"
  ) %>%
  # create a sgender column based on y1/y2
  mutate(
    gndr.c = case_when(
      y == "means_y1" ~ -0.5,
      y == "means_y2" ~ 0.5
    )
  ) %>%
  dplyr::select(cntry,gggi.z.cm,means,gndr.c)


ddsc_mod2_GGGI_ri<-
  ddsc_ml(data=data.frame(d_GGGI_long),predictor = "gggi.z.cm",
          moderator = "gndr.c",
        DV = "means",lvl2_unit = "cntry",
        moderator_values = c(-0.5,0.5))
round(ddsc_mod2_GGGI_ri$results,3)

# country-time multilevel model


mod2_GGGI_cntry_year<-
  lmer(tra.z.wt~gndr.c+gggi.z.cm+gndr.c:gggi.z.cm+
                 (gndr.c|cntry),data=diff_dat_cntry_year,REML=F,
               control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_GGGI_cntry_year)
getFE(mod2_GGGI_cntry_year,round=3)
getVC(mod2_GGGI_cntry_year)
r2mlm(mod2_GGGI,bargraph = F)

ddsc_mod2_GGGI_cntry_year<-
  ddsc_ml(model = mod2_GGGI_cntry_year,
          predictor = "gggi.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)

round(ddsc_mod2_GGGI_cntry_year$vpc_at_moderator_values,3)
round(ddsc_mod2_GGGI_cntry_year$descriptives,3)
round(ddsc_mod2_GGGI_cntry_year$results,3)
round(ddsc_mod2_GGGI_cntry_year$re_cov_test,3)

```

### Comparisons across models

```{r}
# full multilevel model (individuals as entries)
round(ddsc_mod2_GGGI$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# country-level structural path model (country-gender means as entries)
round(ddsc_sem_GGGI$results[c("r_xy1","r_xy2","b_11","b_21","q_b11_b21","diff_b11_b21"),],4)
# multilevel model (country-gender means as entries)
round(ddsc_mod2_GGGI_ri$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# multilevel model (country-year-gender means as entries)
round(ddsc_mod2_GGGI_cntry_year$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)

```

### Bootstrap and equivalence test

Takes a lot of time

```{r}

t1<-Sys.time()
mod2_GGGI_booted_fixef <-
  lme4::bootMer(
    x = mod2_GGGI,
    FUN = lme4::fixef,
    nsim = 1000,
    use.u = FALSE,
    seed = 12345,
    type = c("parametric"),
    verbose = FALSE
  )
t2<-Sys.time()
t2-t1
```


```{r}
# obtain all the bootstrap estimates
mod2_GGGI_boot_est <- data.frame(mod2_GGGI_booted_fixef$t)

# calculate estimates
mod2_GGGI_boot_est$w11<-mod2_GGGI_boot_est$gggi.z.cm+(-0.5)*mod2_GGGI_boot_est$gndr.c.gggi.z.cm
mod2_GGGI_boot_est$w21<-mod2_GGGI_boot_est$gggi.z.cm+(0.5)*mod2_GGGI_boot_est$gndr.c.gggi.z.cm
mod2_GGGI_boot_est$b11<-mod2_GGGI_boot_est$w11/ddsc_mod2_GGGI$SDs["SD_pooled"]
mod2_GGGI_boot_est$b21<-mod2_GGGI_boot_est$w21/ddsc_mod2_GGGI$SDs["SD_pooled"]
mod2_GGGI_boot_est$r_xy1<-mod2_GGGI_boot_est$w11/ddsc_mod2_GGGI$SDs["SD_y1"]
mod2_GGGI_boot_est$r_xy2<-mod2_GGGI_boot_est$w21/ddsc_mod2_GGGI$SDs["SD_y2"]
mod2_GGGI_boot_est$q_b<-atanh(mod2_GGGI_boot_est$b11)-atanh(mod2_GGGI_boot_est$b21)
mod2_GGGI_boot_est$q<-atanh(mod2_GGGI_boot_est$r_xy1)-atanh(mod2_GGGI_boot_est$r_xy2)

# Calculate bootstrap summary statistics
mod2_GGGI_boot_results <- t(as.data.frame(sapply(
  mod2_GGGI_boot_est,
  function(x) {
    c(
      Estimate = mean(x, na.rm = TRUE),
      SE = stats::sd(x, na.rm = TRUE),
      stats::quantile(x, c((1 - .95) / 2,
                           1 - (1 - .95) / 2), na.rm = TRUE)
    )
  }
)))

mod2_GGGI_boot_results

# equivalence test for q_b
tost_z(est=mod2_GGGI_boot_results["q_b","Estimate"],
       se=mod2_GGGI_boot_results["q_b","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

# equivalence test for q
tost_z(est=mod2_GGGI_boot_results["q","Estimate"],
       se=mod2_GGGI_boot_results["q","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

```



### Figure

```{r}
# plot the results

# start with obtaining predicted values for means and differences

# refit reduced and full models with GGGI in original scale


mod2_GGGI_unstd<-lmer(tra.z~gndr.c+gggi.cm+gndr.c:gggi.cm+
                       (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
                     control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))

mod2_GGGI_unstd_red<-lmer(tra.z~gndr.c+
                           (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
                         control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))


p<-
  emmip(
    mod2_GGGI_unstd, 
    gndr.c ~ gggi.cm,
    at=list(gndr.c = c(-0.5,0.5),
            gggi.cm=
              seq(from=round(range(GGGI_country_means$gggi.cm,na.rm=T)[1],2),
                  to=round(range(GGGI_country_means$gggi.cm,na.rm=T)[2],2),
                  by=0.001)),
    plotit=F,CIs=T,lmerTest.limit = 1e6,disable.pbkrtest=T)

p$gndr.c<-p$tvar
levels(p$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.comp<-min(p$LCL)
max.y.comp<-max(p$UCL)

# Men and Women mean distributions

p3<-coefficients(mod2_GGGI_unstd_red)$cntry
p3<-cbind(rbind(p3,p3),weight=rep(c(-0.5,0.5),each=nrow(p3)))
p3$xvar<-p3$`(Intercept)`+p3$gndr.c*p3$weight
p3$gndr.c<-as.factor(p3$weight)
levels(p3$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.mean.distr<-min(p3$xvar)
max.y.mean.distr<-max(p3$xvar)


# obtain the coefs for the gndr.c-effect (difference) as function of gggi.cm

p2<-data.frame(
  emtrends(mod2_GGGI_unstd,var="gndr.c",
           specs="gggi.cm",
           at=list(#gndr.c = c(-0.5,0.5),
             gggi.cm=
               seq(from=round(range(GGGI_country_means$gggi.cm,na.rm=T)[1],2),
                   to=round(range(GGGI_country_means$gggi.cm,na.rm=T)[2],2),
                   by=0.001)),
           lmerTest.limit = 1e6,disable.pbkrtest=T))

p2$yvar<-p2$gndr.c.trend
p2$xvar<-p2$gggi.cm
p2$LCL<-p2$lower.CL
p2$UCL<-p2$upper.CL

# obtain min and max for aligned plots
min.y.diff<-min(p2$LCL)
max.y.diff<-max(p2$UCL)

# difference score distribution

p4<-coefficients(mod2_GGGI_unstd_red)$cntry
p4$xvar=(+1)*p4$gndr.c

# obtain mix and max for aligned plots

min.y.diff.distr<-min(p4$xvar)
max.y.diff.distr<-max(p4$xvar)

# define mins and maxs

min.y.pred<-
  ifelse(min.y.comp<min.y.mean.distr,min.y.comp,min.y.mean.distr)

max.y.pred<-
  ifelse(max.y.comp>max.y.mean.distr,max.y.comp,max.y.mean.distr)

min.y.narrow<-
  ifelse(min.y.diff<min.y.diff.distr,min.y.diff,min.y.diff.distr)

max.y.narrow<-
  ifelse(max.y.diff>max.y.diff.distr,max.y.diff,max.y.diff.distr)

# Figures 

# p1

# scaled simple effects to the plot
pvals<-round_tidy(ddsc_mod2_GGGI$results[6:7,"p.value"],3)

ests<-
  round_tidy(ddsc_mod2_GGGI$results[6:7,"estimate"],2)

coef1<-paste0("std. b11 = ",ests[1],", p = ",pvals[1])
coef2<-paste0("std. b21 = ",ests[2],", p = ",pvals[2])
coefs<-data.frame(gndr.c=c("Women","Men"),
                  coef=c(coef1,coef2))

coef_q<-paste0("Cohen's q = ",round_tidy(ddsc_mod2_GGGI$results["q_b11_b21","estimate"],2),", p = ",
               substr(round_tidy(ddsc_mod2_GGGI$results["interaction","p.value"],3),2,5))  
# prediction plot for difference score

pvals2<-round_tidy(ddsc_mod2_GGGI$results["interaction","p.value"],3)

ests2<-
  round_tidy(-1*ddsc_mod2_GGGI$results["r_xy1y2","estimate"],2)

coefs2<-paste0("difference score correlation = ",ests2,", p = ",pvals2)
# attempt to make a flag-plot

flag_points_GGGI<-coefficients(mod2_GGGI_unstd_red)$cntry

flag_points_GGGI$women<-flag_points_GGGI$`(Intercept)`+(-0.5)*flag_points_GGGI$gndr.c
flag_points_GGGI$men<-flag_points_GGGI$`(Intercept)`+(0.5)*flag_points_GGGI$gndr.c
flag_points_GGGI$mean_level<-flag_points_GGGI$`(Intercept)`
flag_points_GGGI$difference<-flag_points_GGGI$men-flag_points_GGGI$women
flag_points_GGGI$cntry<-rownames(flag_points_GGGI)

flag_points_GGGI<-
  left_join(x=flag_points_GGGI,
            y=GGGI_country_means[,c("ISO2","gggi.cm")],by=c("cntry"="ISO2"))
#flag_points_GGGI

flag_points_GGGI_long<-
  data.frame(mean_level=c(flag_points_GGGI$women,
                          flag_points_GGGI$men),
             gggi.cm=c(flag_points_GGGI$gggi.cm,
                                 flag_points_GGGI$gggi.cm),
             cntry=c(flag_points_GGGI$cntry,
                     flag_points_GGGI$cntry),
             gndr.c=rep(c("Women","Men"),each=nrow(flag_points_GGGI)
             ))


p1.tra.flags<-
  ggplot(p,aes(y=yvar,x=gggi.cm,color=gndr.c))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("Global Gender Gap Index (Average 2002-2023)")+
  #ylim(c(min.y.pred,max.y.pred))+
  #xlim(c(0.60,1.00))+
  ylab("Value tradition (Average 2002-2023)")+
  scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "top",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(linewidth = 0.5, linetype = 2,
                                          colour = "gray"))+
  geom_text(data = coefs,show.legend=F,
            aes(label=coef,x=0.61,
                y=c(-0.35,-0.40),size=14,hjust="left"))+
  geom_text(inherit.aes=F,aes(x=0.61,y=-0.45,
                              label=coef_q,size=14,hjust="left"),
            show.legend=F)+
  geom_point(data=flag_points_GGGI_long,size=8,alpha=0.30,
             aes(x=gggi.cm,y=mean_level))+
  geom_line(data=flag_points_GGGI_long,aes(group = cntry,x=gggi.cm,y=mean_level),
            color="black",linetype=2)+
  geom_flag(data=flag_points_GGGI_long,show.legend=F,
            aes(country=tolower(cntry),size=14,x=gggi.cm,y=mean_level))

p2.tra.flags<-ggplot(p2,aes(y=yvar,x=gggi.cm))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("Global Gender Gap Index (Average 2002-2023)")+
  #xlim(c(0.60,1.00))+
  #ylim(c(min.y.narrow,max.y.narrow))+
  ylab("Difference in value tradition")+
  #scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "right",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(linewidth = 0.5, linetype = 2,
                                          colour = "gray"))+
  #geom_text(coef2,aes(x=0.63,y=min(p2$LCL)))
  geom_text(data = data.frame(coefs2),show.legend=F,
            aes(label=coefs2,x=0.70,
                y=c(round(min(p2$LCL),2))+0.05,size=18,hjust="left"))+
  geom_flag(data=flag_points_GGGI,show.legend=F,
            aes(country=tolower(cntry),size=18,x=gggi.cm,y=difference))

pflag_comb<-
  ggarrange(p1.tra.flags,p2.tra.flags,align = "v",
            ncol=1,nrow=2,heights=c(2,1))
pflag_comb


png(filename = 
      "../results/tra/GGGI_flags_new.png",
    units = "cm",
    width = 21.0,height=29.7*(3/4),res = 300)
pflag_comb
dev.off()
```


## mod2 with Gender-equality index (GDI)

```{r}
mod2_GDI<-lmer(tra.z~gndr.c+gdi.z.cm+gndr.c:gdi.z.cm+
                 (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_GDI)
getFE(mod2_GDI,round=3)
getVC(mod2_GDI)
r2mlm(mod2_GDI,bargraph = F)
```


### Deconstructed associations

```{r}
t1<-Sys.time()
ddsc_mod2_GDI<-
  ddsc_ml(model = mod2_GDI,
          predictor = "gdi.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)
t2<-Sys.time()
t2-t1
round(ddsc_mod2_GDI$vpc_at_moderator_values,3)
round(ddsc_mod2_GDI$descriptives,3)
round(ddsc_mod2_GDI$results,3)
round(ddsc_mod2_GDI$re_cov_test,3)

# Structural path model
d_GDI<-ddsc_mod2_GDI$ddsc_sem_fit$data

ddsc_sem_GDI<-
  ddsc_sem(data=d_GDI,x = "gdi.z.cm",y1="means_y2",
           y2="means_y1",q_sesoi = .10)
round(ddsc_sem_GDI$results,3)
round(ddsc_sem_GDI$variance_test,3)

## random intercept mlm with double-entries for each country (men and women)

d_GDI_long <- d_GDI %>%
  # move row names into a column
  rownames_to_column("cntry") %>%
  # pivot only the means_y1 / means_y2 columns
  pivot_longer(
    cols = c(means_y1, means_y2),
    names_to = "y",
    values_to = "means"
  ) %>%
  # create a sgender column based on y1/y2
  mutate(
    gndr.c = case_when(
      y == "means_y1" ~ -0.5,
      y == "means_y2" ~ 0.5
    )
  ) %>%
  dplyr::select(cntry,gdi.z.cm,means,gndr.c)


ddsc_mod2_GDI_ri<-
  ddsc_ml(data=data.frame(d_GDI_long),predictor = "gdi.z.cm",
          moderator = "gndr.c",
        DV = "means",lvl2_unit = "cntry",
        moderator_values = c(-0.5,0.5))
round(ddsc_mod2_GDI_ri$results,3)

# country-time multilevel model


mod2_GDI_cntry_year<-
  lmer(tra.z.wt~gndr.c+gdi.z.cm+gndr.c:gdi.z.cm+
                 (gndr.c|cntry),data=diff_dat_cntry_year,REML=F,
               control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_GDI_cntry_year)
getFE(mod2_GDI_cntry_year,round=3)
getVC(mod2_GDI_cntry_year)
r2mlm(mod2_GDI,bargraph = F)

ddsc_mod2_GDI_cntry_year<-
  ddsc_ml(model = mod2_GDI_cntry_year,
          predictor = "gdi.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)

round(ddsc_mod2_GDI_cntry_year$vpc_at_moderator_values,3)
round(ddsc_mod2_GDI_cntry_year$descriptives,3)
round(ddsc_mod2_GDI_cntry_year$results,3)
round(ddsc_mod2_GDI_cntry_year$re_cov_test,3)

```

### Comparisons across models

```{r}
# full multilevel model (individuals as entries)
round(ddsc_mod2_GDI$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# country-level structural path model (country-gender means as entries)
round(ddsc_sem_GDI$results[c("r_xy1","r_xy2","b_11","b_21","q_b11_b21","diff_b11_b21"),],4)
# multilevel model (country-gender means as entries)
round(ddsc_mod2_GDI_ri$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# multilevel model (country-year-gender means as entries)
round(ddsc_mod2_GDI_cntry_year$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)

```

### Bootstrap and equivalence test

Takes a lot of time

```{r}

t1<-Sys.time()
mod2_GDI_booted_fixef <-
  lme4::bootMer(
    x = mod2_GDI,
    FUN = lme4::fixef,
    nsim = 1000,
    use.u = FALSE,
    seed = 12345,
    type = c("parametric"),
    verbose = FALSE
  )
t2<-Sys.time()
t2-t1
```


```{r}
# obtain all the bootstrap estimates
mod2_GDI_boot_est <- data.frame(mod2_GDI_booted_fixef$t)

# calculate estimates
mod2_GDI_boot_est$w11<-mod2_GDI_boot_est$gdi.z.cm+(-0.5)*mod2_GDI_boot_est$gndr.c.gdi.z.cm
mod2_GDI_boot_est$w21<-mod2_GDI_boot_est$gdi.z.cm+(0.5)*mod2_GDI_boot_est$gndr.c.gdi.z.cm
mod2_GDI_boot_est$b11<-mod2_GDI_boot_est$w11/ddsc_mod2_GDI$SDs["SD_pooled"]
mod2_GDI_boot_est$b21<-mod2_GDI_boot_est$w21/ddsc_mod2_GDI$SDs["SD_pooled"]
mod2_GDI_boot_est$r_xy1<-mod2_GDI_boot_est$w11/ddsc_mod2_GDI$SDs["SD_y1"]
mod2_GDI_boot_est$r_xy2<-mod2_GDI_boot_est$w21/ddsc_mod2_GDI$SDs["SD_y2"]
mod2_GDI_boot_est$q_b<-atanh(mod2_GDI_boot_est$b11)-atanh(mod2_GDI_boot_est$b21)
mod2_GDI_boot_est$q<-atanh(mod2_GDI_boot_est$r_xy1)-atanh(mod2_GDI_boot_est$r_xy2)

# Calculate bootstrap summary statistics
mod2_GDI_boot_results <- t(as.data.frame(sapply(
  mod2_GDI_boot_est,
  function(x) {
    c(
      Estimate = mean(x, na.rm = TRUE),
      SE = stats::sd(x, na.rm = TRUE),
      stats::quantile(x, c((1 - .95) / 2,
                           1 - (1 - .95) / 2), na.rm = TRUE)
    )
  }
)))

mod2_GDI_boot_results

# equivalence test for q_b
tost_z(est=mod2_GDI_boot_results["q_b","Estimate"],
       se=mod2_GDI_boot_results["q_b","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

# equivalence test for q
tost_z(est=mod2_GDI_boot_results["q","Estimate"],
       se=mod2_GDI_boot_results["q","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

```



### Figure

```{r}
# plot the results

# start with obtaining predicted values for means and differences


mod2_GDI_unstd<-lmer(tra.z~gndr.c+gdi.cm+gndr.c:gdi.cm+
                       (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
                     control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))

mod2_GDI_unstd_red<-lmer(tra.z~gndr.c+
                           (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
                         control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))


p<-
  emmip(
    mod2_GDI_unstd, 
    gndr.c ~ gdi.cm,
    at=list(gndr.c = c(-0.5,0.5),
            gdi.cm=
              seq(from=round(range(GDI_country_means$gdi.cm,na.rm=T)[1],2),
                  to=round(range(GDI_country_means$gdi.cm,na.rm=T)[2],2),
                  by=0.001)),
    plotit=F,CIs=T,lmerTest.limit = 1e6,disable.pbkrtest=T)

p$gndr.c<-p$tvar
levels(p$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.comp<-min(p$LCL)
max.y.comp<-max(p$UCL)

# Men and Women mean distributions

p3<-coefficients(mod2_GDI_unstd_red)$cntry
p3<-cbind(rbind(p3,p3),weight=rep(c(-0.5,0.5),each=nrow(p3)))
p3$xvar<-p3$`(Intercept)`+p3$gndr.c*p3$weight
p3$gndr.c<-as.factor(p3$weight)
levels(p3$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.mean.distr<-min(p3$xvar)
max.y.mean.distr<-max(p3$xvar)


# obtain the coefs for the gndr.c-effect (difference) as function of gdi.cm

p2<-data.frame(
  emtrends(mod2_GDI_unstd,var="gndr.c",
           specs="gdi.cm",
           at=list(#gndr.c = c(-0.5,0.5),
             gdi.cm=
               seq(from=round(range(GDI_country_means$gdi.cm,na.rm=T)[1],2),
                   to=round(range(GDI_country_means$gdi.cm,na.rm=T)[2],2),
                   by=0.001)),
           lmerTest.limit = 1e6,disable.pbkrtest=T))

p2$yvar<-p2$gndr.c.trend
p2$xvar<-p2$gdi.cm
p2$LCL<-p2$lower.CL
p2$UCL<-p2$upper.CL

# obtain min and max for aligned plots
min.y.diff<-min(p2$LCL)
max.y.diff<-max(p2$UCL)

# difference score distribution

p4<-coefficients(mod2_GDI_unstd_red)$cntry
p4$xvar=(+1)*p4$gndr.c

# obtain mix and max for aligned plots

min.y.diff.distr<-min(p4$xvar)
max.y.diff.distr<-max(p4$xvar)

# define mins and maxs

min.y.pred<-
  ifelse(min.y.comp<min.y.mean.distr,min.y.comp,min.y.mean.distr)

max.y.pred<-
  ifelse(max.y.comp>max.y.mean.distr,max.y.comp,max.y.mean.distr)

min.y.narrow<-
  ifelse(min.y.diff<min.y.diff.distr,min.y.diff,min.y.diff.distr)

max.y.narrow<-
  ifelse(max.y.diff>max.y.diff.distr,max.y.diff,max.y.diff.distr)

# Figures 

# p1

# scaled simple effects to the plot
pvals<-round_tidy(ddsc_mod2_GDI$results[6:7,"p.value"],3)

ests<-
  round_tidy(ddsc_mod2_GDI$results[6:7,"estimate"],2)

coef1<-paste0("std. b11 = ",ests[1],", p = ",pvals[1])
coef2<-paste0("std. b21 = ",ests[2],", p = ",pvals[2])
coefs<-data.frame(gndr.c=c("Women","Men"),
                  coef=c(coef1,coef2))

coef_q<-paste0("Cohen's q = ",round_tidy(ddsc_mod2_GDI$results["q_b11_b21","estimate"],2),", p = ",
               substr(round_tidy(ddsc_mod2_GDI$results["interaction","p.value"],3),2,5))  
# prediction plot for difference score

pvals2<-round_tidy(ddsc_mod2_GDI$results["interaction","p.value"],3)

ests2<-
  round_tidy(-1*ddsc_mod2_GDI$results["r_xy1y2","estimate"],2)

coefs2<-paste0("difference score correlation = ",ests2,", p = ",pvals2)
# attempt to make a flag-plot

flag_points_GDI<-coefficients(mod2_GDI_unstd_red)$cntry

flag_points_GDI$women<-flag_points_GDI$`(Intercept)`+(-0.5)*flag_points_GDI$gndr.c
flag_points_GDI$men<-flag_points_GDI$`(Intercept)`+(0.5)*flag_points_GDI$gndr.c
flag_points_GDI$mean_level<-flag_points_GDI$`(Intercept)`
flag_points_GDI$difference<-flag_points_GDI$men-flag_points_GDI$women
flag_points_GDI$cntry<-rownames(flag_points_GDI)

flag_points_GDI<-
  left_join(x=flag_points_GDI,
            y=GDI_country_means[,c("ISO2","gdi.cm")],by=c("cntry"="ISO2"))
#flag_points_GDI

flag_points_GDI_long<-
  data.frame(mean_level=c(flag_points_GDI$women,
                          flag_points_GDI$men),
             gdi.cm=c(flag_points_GDI$gdi.cm,
                                 flag_points_GDI$gdi.cm),
             cntry=c(flag_points_GDI$cntry,
                     flag_points_GDI$cntry),
             gndr.c=rep(c("Women","Men"),each=nrow(flag_points_GDI)
             ))


p1.tra.flags<-
  ggplot(p,aes(y=yvar,x=gdi.cm,color=gndr.c))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("Gender Development Index (Average 2002-2023)")+
  #ylim(c(min.y.pred,max.y.pred))+
  #xlim(c(0.60,1.00))+
  ylab("Value tradition (Average 2002-2023)")+
  scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "top",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(linewidth = 0.5, linetype = 2,
                                          colour = "gray"))+
  geom_text(data = coefs,show.legend=F,
            aes(label=coef,x=0.90,
                y=c(-0.35,-0.40),size=14,hjust="left"))+
  geom_text(inherit.aes=F,aes(x=0.90,y=-0.45,
                              label=coef_q,size=14,hjust="left"),
            show.legend=F)+
  geom_point(data=flag_points_GDI_long,size=8,alpha=0.30,
             aes(x=gdi.cm,y=mean_level))+
  geom_line(data=flag_points_GDI_long,aes(group = cntry,x=gdi.cm,y=mean_level),
            color="black",linetype=2)+
  geom_flag(data=flag_points_GDI_long,show.legend=F,
            aes(country=tolower(cntry),size=14,x=gdi.cm,y=mean_level))

#p1.tra.flags


p2.tra.flags<-ggplot(p2,aes(y=yvar,x=gdi.cm))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("Gender Development Index (Average 2002-2023)")+
  #xlim(c(0.60,1.00))+
  #ylim(c(min.y.narrow,max.y.narrow))+
  ylab("Difference in value tradition")+
  #scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "right",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(size = 0.5, linetype = 2,
                                          colour = "gray"))+
  #geom_text(coef2,aes(x=0.63,y=min(p2$LCL)))
  geom_text(data = data.frame(coefs2),show.legend=F,
            aes(label=coefs2,x=0.96,
                y=c(round(min(p2$LCL),2)),size=18,hjust="left"))+
  geom_flag(data=flag_points_GDI,show.legend=F,
            aes(country=tolower(cntry),size=18,x=gdi.cm,y=difference))

#p2.tra.flags


pflag_comb<-
  ggarrange(p1.tra.flags,p2.tra.flags,align = "v",
            ncol=1,nrow=2,heights=c(2,1))
pflag_comb


png(filename = 
      "../results/tra/GDI_flags.png",
    units = "cm",
    width = 21.0,height=29.7*(3/4),res = 300)
pflag_comb
dev.off()
```


## mod2 with Gross Domestic Product (log_GDP)

* Logarithm of GDP per capita, PPP (constant 2017 international $)

```{r}
mod2_log_GDP<-lmer(tra.z~gndr.c+log_gdp.z.cm+
                     gndr.c:log_gdp.z.cm+
                 (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_log_GDP)
getFE(mod2_log_GDP,round=3)
getVC(mod2_log_GDP)
r2mlm(mod2_log_GDP,bargraph = F)
```


### Deconstructed associations

```{r}
t1<-Sys.time()
ddsc_mod2_log_GDP<-
  ddsc_ml(model = mod2_log_GDP,
          predictor = "log_gdp.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)
t2<-Sys.time()
t2-t1
round(ddsc_mod2_log_GDP$vpc_at_moderator_values,3)
round(ddsc_mod2_log_GDP$descriptives,3)
round(ddsc_mod2_log_GDP$results,3)
round(ddsc_mod2_log_GDP$re_cov_test,3)

# Structural path model
d_log_GDP<-ddsc_mod2_log_GDP$ddsc_sem_fit$data

ddsc_sem_log_GDP<-
  ddsc_sem(data=d_log_GDP,x = "log_gdp.z.cm",y1="means_y2",
           y2="means_y1",q_sesoi = .10)
round(ddsc_sem_log_GDP$results,3)
round(ddsc_sem_log_GDP$variance_test,3)

## random intercept mlm with double-entries for each country (men and women)

d_log_GDP_long <- d_log_GDP %>%
  # move row names into a column
  rownames_to_column("cntry") %>%
  # pivot only the means_y1 / means_y2 columns
  pivot_longer(
    cols = c(means_y1, means_y2),
    names_to = "y",
    values_to = "means"
  ) %>%
  # create a sgender column based on y1/y2
  mutate(
    gndr.c = case_when(
      y == "means_y1" ~ -0.5,
      y == "means_y2" ~ 0.5
    )
  ) %>%
  dplyr::select(cntry,log_gdp.z.cm,means,gndr.c)


ddsc_mod2_log_GDP_ri<-
  ddsc_ml(data=data.frame(d_log_GDP_long),predictor = "log_gdp.z.cm",
          moderator = "gndr.c",
        DV = "means",lvl2_unit = "cntry",
        moderator_values = c(-0.5,0.5))
round(ddsc_mod2_log_GDP_ri$results,3)

# country-time multilevel model


mod2_log_GDP_cntry_year<-
  lmer(tra.z.wt~gndr.c+log_gdp.z.cm+gndr.c:log_gdp.z.cm+
                 (gndr.c|cntry),data=diff_dat_cntry_year,REML=F,
               control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod2_log_GDP_cntry_year)
getFE(mod2_log_GDP_cntry_year,round=3)
getVC(mod2_log_GDP_cntry_year)
r2mlm(mod2_log_GDP,bargraph = F)

ddsc_mod2_log_GDP_cntry_year<-
  ddsc_ml(model = mod2_log_GDP_cntry_year,
          predictor = "log_gdp.z.cm",
          moderator = "gndr.c",moderator_values = c(-0.5,0.5),
          re_cov_test = T)

round(ddsc_mod2_log_GDP_cntry_year$vpc_at_moderator_values,3)
round(ddsc_mod2_log_GDP_cntry_year$descriptives,3)
round(ddsc_mod2_log_GDP_cntry_year$results,3)
round(ddsc_mod2_log_GDP_cntry_year$re_cov_test,3)

```

### Comparisons across models

```{r}
# full multilevel model (individuals as entries)
round(ddsc_mod2_log_GDP$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# country-level structural path model (country-gender means as entries)
round(ddsc_sem_log_GDP$results[c("r_xy1","r_xy2","b_11","b_21","q_b11_b21","diff_b11_b21"),],4)
# multilevel model (country-gender means as entries)
round(ddsc_mod2_log_GDP_ri$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)
# multilevel model (country-year-gender means as entries)
round(ddsc_mod2_log_GDP_cntry_year$results[c("r_xy1","r_xy2","b_11","b_21","main_effect","moderator_effect","interaction","q_b11_b21"),],4)

```

### Bootstrap and equivalence test

Takes a lot of time

```{r}

t1<-Sys.time()
mod2_log_GDP_booted_fixef <-
  lme4::bootMer(
    x = mod2_log_GDP,
    FUN = lme4::fixef,
    nsim = 1000,
    use.u = FALSE,
    seed = 12345,
    type = c("parametric"),
    verbose = FALSE
  )
t2<-Sys.time()
t2-t1
```


```{r}
# obtain all the bootstrap estimates
mod2_log_GDP_boot_est <- data.frame(mod2_log_GDP_booted_fixef$t)

# calculate estimates
mod2_log_GDP_boot_est$w11<-mod2_log_GDP_boot_est$log_gdp.z.cm+(-0.5)*mod2_log_GDP_boot_est$gndr.c.log_gdp.z.cm
mod2_log_GDP_boot_est$w21<-mod2_log_GDP_boot_est$log_gdp.z.cm+(0.5)*mod2_log_GDP_boot_est$gndr.c.log_gdp.z.cm
mod2_log_GDP_boot_est$b11<-mod2_log_GDP_boot_est$w11/ddsc_mod2_log_GDP$SDs["SD_pooled"]
mod2_log_GDP_boot_est$b21<-mod2_log_GDP_boot_est$w21/ddsc_mod2_log_GDP$SDs["SD_pooled"]
mod2_log_GDP_boot_est$r_xy1<-mod2_log_GDP_boot_est$w11/ddsc_mod2_log_GDP$SDs["SD_y1"]
mod2_log_GDP_boot_est$r_xy2<-mod2_log_GDP_boot_est$w21/ddsc_mod2_log_GDP$SDs["SD_y2"]
mod2_log_GDP_boot_est$q_b<-atanh(mod2_log_GDP_boot_est$b11)-atanh(mod2_log_GDP_boot_est$b21)
mod2_log_GDP_boot_est$q<-atanh(mod2_log_GDP_boot_est$r_xy1)-atanh(mod2_log_GDP_boot_est$r_xy2)

# Calculate bootstrap summary statistics
mod2_log_GDP_boot_results <- t(as.data.frame(sapply(
  mod2_log_GDP_boot_est,
  function(x) {
    c(
      Estimate = mean(x, na.rm = TRUE),
      SE = stats::sd(x, na.rm = TRUE),
      stats::quantile(x, c((1 - .95) / 2,
                           1 - (1 - .95) / 2), na.rm = TRUE)
    )
  }
)))

mod2_log_GDP_boot_results

# equivalence test for q_b
tost_z(est=mod2_log_GDP_boot_results["q_b","Estimate"],
       se=mod2_log_GDP_boot_results["q_b","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

# equivalence test for q
tost_z(est=mod2_log_GDP_boot_results["q","Estimate"],
       se=mod2_log_GDP_boot_results["q","SE"],
       low=-.10,
       high=.10, alpha = 0.05)

```



### Figure

```{r}
# plot the results

# start with obtaining predicted values for means and differences


mod2_log_GDP_unstd<-lmer(tra.z~gndr.c+log_gdp.cm+
                           gndr.c:log_gdp.cm+
                       (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
                     control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))

mod2_log_GDP_unstd_red<-lmer(tra.z~gndr.c+
                           (gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
                         control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))


p<-
  emmip(
    mod2_log_GDP_unstd, 
    gndr.c ~ log_gdp.cm,
    at=list(gndr.c = c(-0.5,0.5),
            log_gdp.cm=
              seq(from=round(range(GDP_country_means$log_gdp.cm,na.rm=T)[1],2),
                  to=round(range(GDP_country_means$log_gdp.cm,na.rm=T)[2],2),
                  by=0.001)),
    plotit=F,CIs=T,lmerTest.limit = 1e6,disable.pbkrtest=T)

p$gndr.c<-p$tvar
levels(p$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.comp<-min(p$LCL)
max.y.comp<-max(p$UCL)

# Men and Women mean distributions

p3<-coefficients(mod2_log_GDP_unstd_red)$cntry
p3<-cbind(rbind(p3,p3),weight=rep(c(-0.5,0.5),each=nrow(p3)))
p3$xvar<-p3$`(Intercept)`+p3$gndr.c*p3$weight
p3$gndr.c<-as.factor(p3$weight)
levels(p3$gndr.c)<-c("Women","Men")

# obtain min and max for aligned plots
min.y.mean.distr<-min(p3$xvar)
max.y.mean.distr<-max(p3$xvar)


# obtain the coefs for the gndr.c-effect (difference) as function of log_gdp.cm

p2<-data.frame(
  emtrends(mod2_log_GDP_unstd,var="gndr.c",
           specs="log_gdp.cm",
           at=list(#gndr.c = c(-0.5,0.5),
             log_gdp.cm=
               seq(from=round(range(GDP_country_means$log_gdp.cm,na.rm=T)[1],2),
                   to=round(range(GDP_country_means$log_gdp.cm,na.rm=T)[2],2),
                   by=0.001)),
           lmerTest.limit = 1e6,disable.pbkrtest=T))

p2$yvar<-p2$gndr.c.trend
p2$xvar<-p2$log_gdp.cm
p2$LCL<-p2$lower.CL
p2$UCL<-p2$upper.CL

# obtain min and max for aligned plots
min.y.diff<-min(p2$LCL)
max.y.diff<-max(p2$UCL)

# difference score distribution

p4<-coefficients(mod2_log_GDP_unstd_red)$cntry
p4$xvar=(+1)*p4$gndr.c

# obtain mix and max for aligned plots

min.y.diff.distr<-min(p4$xvar)
max.y.diff.distr<-max(p4$xvar)

# define mins and maxs

min.y.pred<-
  ifelse(min.y.comp<min.y.mean.distr,min.y.comp,min.y.mean.distr)

max.y.pred<-
  ifelse(max.y.comp>max.y.mean.distr,max.y.comp,max.y.mean.distr)

min.y.narrow<-
  ifelse(min.y.diff<min.y.diff.distr,min.y.diff,min.y.diff.distr)

max.y.narrow<-
  ifelse(max.y.diff>max.y.diff.distr,max.y.diff,max.y.diff.distr)

# Figures 

# p1

# scaled simple effects to the plot
pvals<-round_tidy(ddsc_mod2_log_GDP$results[6:7,"p.value"],3)

ests<-
  round_tidy(ddsc_mod2_log_GDP$results[6:7,"estimate"],2)

coef1<-paste0("std. b11 = ",ests[1],", p = ",pvals[1])
coef2<-paste0("std. b21 = ",ests[2],", p = ",pvals[2])
coefs<-data.frame(gndr.c=c("Women","Men"),
                  coef=c(coef1,coef2))

coef_q<-paste0("Cohen's q = ",round_tidy(ddsc_mod2_log_GDP$results["q_b11_b21","estimate"],2),", p = ",
               substr(round_tidy(ddsc_mod2_log_GDP$results["interaction","p.value"],3),2,5))  
# prediction plot for difference score

pvals2<-round_tidy(ddsc_mod2_log_GDP$results["interaction","p.value"],3)

ests2<-
  round_tidy(-1*ddsc_mod2_log_GDP$results["r_xy1y2","estimate"],2)

coefs2<-paste0("difference score correlation = ",ests2,", p = ",pvals2)
# attempt to make a flag-plot

flag_points_log_GDP<-coefficients(mod2_log_GDP_unstd_red)$cntry

flag_points_log_GDP$women<-flag_points_log_GDP$`(Intercept)`+(-0.5)*flag_points_log_GDP$gndr.c
flag_points_log_GDP$men<-flag_points_log_GDP$`(Intercept)`+(0.5)*flag_points_log_GDP$gndr.c
flag_points_log_GDP$mean_level<-flag_points_log_GDP$`(Intercept)`
flag_points_log_GDP$difference<-flag_points_log_GDP$men-flag_points_log_GDP$women
flag_points_log_GDP$cntry<-rownames(flag_points_log_GDP)

flag_points_log_GDP<-
  left_join(x=flag_points_log_GDP,
            y=GDP_country_means[,c("ISO2","log_gdp.cm")],by=c("cntry"="ISO2"))
#flag_points_log_GDP

flag_points_log_GDP_long<-
  data.frame(mean_level=c(flag_points_log_GDP$women,
                          flag_points_log_GDP$men),
             log_gdp.cm=c(flag_points_log_GDP$log_gdp.cm,
                                 flag_points_log_GDP$log_gdp.cm),
             cntry=c(flag_points_log_GDP$cntry,
                     flag_points_log_GDP$cntry),
             gndr.c=rep(c("Women","Men"),each=nrow(flag_points_log_GDP)
             ))


p1.tra.flags<-
  ggplot(p,aes(y=yvar,x=log_gdp.cm,color=gndr.c))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("log(GDP per capita) (Average 2002-2023)")+
  #ylim(c(min.y.pred,max.y.pred))+
  #xlim(c(0.60,1.00))+
  ylab("Value tradition (Average 2002-2023)")+
  scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "top",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(linewidth = 0.5, linetype = 2,
                                          colour = "gray"))+
  geom_text(data = coefs,show.legend=F,
            aes(label=coef,x=9.75,
                y=c(-0.35,-0.40),size=14,hjust="left"))+
  geom_text(inherit.aes=F,aes(x=9.75,y=-0.45,
                              label=coef_q,size=14,hjust="left"),
            show.legend=F)+
  geom_point(data=flag_points_log_GDP_long,size=8,alpha=0.30,
             aes(x=log_gdp.cm,y=mean_level))+
  geom_line(data=flag_points_log_GDP_long,aes(group = cntry,x=log_gdp.cm,y=mean_level),
            color="black",linetype=2)+
  geom_flag(data=flag_points_log_GDP_long,show.legend=F,
            aes(country=tolower(cntry),size=14,x=log_gdp.cm,y=mean_level))

p2.tra.flags<-ggplot(p2,aes(y=yvar,x=log_gdp.cm))+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=LCL, ymax=UCL),alpha=0.2)+
  xlab("log(GDP per capita) (Average 2002-2023)")+
  #xlim(c(0.60,1.00))+
  #ylim(c(min.y.narrow,max.y.narrow))+
  ylab("Difference in value tradition")+
  #scale_color_manual(values=met.brewer("Archambault")[c(6,2)])+
  theme(legend.position = "right",
        legend.title=element_blank(),
        text=element_text(size=16,  family="sans"),
        panel.background = element_rect(fill = "white",
                                        #colour = "black",
                                        #size = 0.5, linetype = "solid"
        ),
        panel.grid.major.x = element_line(size = 0.5, linetype = 2,
                                          colour = "gray"))+
  #geom_text(coef2,aes(x=0.63,y=min(p2$LCL)))
  geom_text(data = data.frame(coefs2),show.legend=F,
            aes(label=coefs2,x=9.75,
                y=c(round(min(p2$LCL),2)),size=18,hjust="left"))+
  geom_flag(data=flag_points_log_GDP,show.legend=F,
            aes(country=tolower(cntry),size=18,x=log_gdp.cm,y=difference))


pflag_comb<-
  ggarrange(p1.tra.flags,p2.tra.flags,align = "v",
            ncol=1,nrow=2,heights=c(2,1))
pflag_comb


png(filename = 
      "../results/tra/log_GDP_flags.png",
    units = "cm",
    width = 21.0,height=29.7*(3/4),res = 300)
pflag_comb
dev.off()
```


## mod3: fixed effect of time (year)

* Year coded year-2002 (2002 is a zero-point)

```{r}
mod3<-lmer(tra.z~gndr.c+essround.c+(gndr.c|cntry),data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod3)
getFE(mod3,round=3)
getVC(mod3)
r2mlm(mod3,bargraph = F)
anova(mod2,mod3)
```

## mod4: random effect of time (year)

```{r}
mod4<-lmer(tra.z~gndr.c+year.c+(gndr.c+year.c|cntry),
           data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod4)
getFE(mod4,round=3)
getVC(mod4)
r2mlm(mod4,bargraph = F)
anova(mod2,mod3,mod4)
```

## mod5: fixed interaction between time and gender

```{r}
mod5<-lmer(tra.z~gndr.c+year.c+
             gndr.c:year.c+(gndr.c+year.c|cntry),
           data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod5)
getFE(mod5,round=3)
getVC(mod5)
r2mlm(mod5,bargraph = F)
anova(mod4,mod5)
```

## mod6: random interaction between time and gender

```{r}
mod6<-lmer(tra.z~gndr.c+year.c+
             gndr.c:year.c+(gndr.c+year.c+gndr.c:year.c|cntry),
           data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod6)
getFE(mod6,round=3)
getVC(mod6)
r2mlm(mod6,bargraph = F)
anova(mod4,mod5,mod6)
```

### Trends

```{r}
# gender specific change over time

change_mod6<-emmeans(mod6,specs="year.c",by="gndr.c",
                     at=list(gndr.c=c(-0.5,0.5),
                             year.c=rev(range(diff_dat$year.c))),
                     disable.pbkrtest=T,
                     lmerTest.limit = 500000,infer=c(T,T))
change_mod6
pairs(change_mod6,adjust="none",infer=c(T,T))

# change in gender differences over time

change_in_diff_mod6<-emmeans(mod6,specs=c("gndr.c","year.c"),
                             at=list(gndr.c=c(-0.5,0.5),
                                     year.c=rev(range(diff_dat$year.c))),
                             disable.pbkrtest=T,
                             lmerTest.limit = 500000,infer=c(T,T))
change_in_diff_mod6
pairs(change_in_diff_mod6,adjust="none",infer=c(T,T))

# custom contrast to obtain the estimates for the entire span

changes_in_diff <- list(
  diff_ESS11 = c(-1,1,0,0),
  diff_ESS1 = c(0,0,-1,1))

diff_mod6<-contrast(change_in_diff_mod6,method = changes_in_diff,adjust="none",infer=c(T,T))

# gender differences at ESS1 and ESS11
diff_mod6
# Test for whether and how much the differences have changed
pairs(diff_mod6,infer=c(T,T))

```

### Equivalence test for differences in linear change

```{r}
diff_in_linear_change<-pairs(diff_mod6,infer=c(T,T))
tost_t(est = data.frame(diff_in_linear_change)[1,"estimate"],
       se = data.frame(diff_in_linear_change)[1,"SE"],
       df = data.frame(diff_in_linear_change)[1,"df"],
       low = -0.20,high = 0.20,alpha = .10)
```

### Figure for time trends

```{r}
# Figure for average patterns

# Model-based development for men and women

p_mod6<-
  emmip(
    mod6, 
    gndr.c ~ year.c,
    at=list(gndr.c = c(-0.5,0.5),
            year.c=
              unique(diff_dat$year.c)),
    plotit=F,CIs=T,lmerTest.limit = 1e6,disable.pbkrtest=T)

p_mod6$gndr.c<-p_mod6$tvar
levels(p_mod6$gndr.c)<-c("Women","Men")

p_mod6<-data.frame(p_mod6)

p_mod6$year<-p_mod6$year.c+2002

# add observed statistics as well

p_mod6$obs_mean<-NA
p_mod6$obs_sd<-NA
p_mod6$obs_n<-NA
p_mod6$obs_mean_wt<-NA
p_mod6$obs_sd_wt<-NA
p_mod6$obs_n_wt<-NA

for(i in 1:nrow(p_mod6)){
  #i=1
  #cntry_i<-p_mod6[i,"cntry"]
  year_i<-p_mod6[i,"year.c"]
  gndr_i<-as.numeric(as.character(p_mod6[i,"tvar"]))
  
  temp_diff_dat<-diff_dat %>%
    filter(#cntry==cntry_i,
      year.c==year_i,
      gndr.c==gndr_i) %>%
    dplyr::summarize(#n=n(),
      obs_n_wt=sum(pspwght),
      obs_mean_wt=weighted.mean(x=tra.z,w=pspwght),
      obs_sd_wt=sqrt(wtd.var(tra.z,pspwght)),
      obs_mean=mean(tra.z),
      obs_sd=sd(tra.z),
      obs_n=n()) 
  
  p_mod6[i,"obs_mean"]<-temp_diff_dat[,"obs_mean"]
  p_mod6[i,"obs_sd"]<-temp_diff_dat[,"obs_sd"]
  p_mod6[i,"obs_n"]<-temp_diff_dat[,"obs_n"]
  p_mod6[i,"obs_mean_wt"]<-temp_diff_dat[,"obs_mean_wt"]
  p_mod6[i,"obs_sd_wt"]<-temp_diff_dat[,"obs_sd_wt"]
  p_mod6[i,"obs_n_wt"]<-temp_diff_dat[,"obs_n_wt"]
}

# calculate confidence intervals
p_mod6$obs_mean_se<-p_mod6$obs_sd/sqrt(p_mod6$obs_n)
p_mod6$obs_mean_LL<-p_mod6$obs_mean+
  qnorm(.025)*p_mod6$obs_mean_se
p_mod6$obs_mean_UL<-p_mod6$obs_mean+
  qnorm(.975)*p_mod6$obs_mean_se

p_mod6$obs_mean_wt_se<-p_mod6$obs_sd_wt/sqrt(p_mod6$obs_n_wt)
p_mod6$obs_mean_wt_LL<-p_mod6$obs_mean_wt+
  qnorm(.025)*p_mod6$obs_mean_wt_se
p_mod6$obs_mean_wt_UL<-p_mod6$obs_mean_wt+
  qnorm(.975)*p_mod6$obs_mean_wt_se


# Figure

#my_colors <- met.brewer("Cassatt2")[c(3, 8)]
my_colors <- met.brewer("Archambault")[c(6,2)]

p_time_trends<-
  ggplot(p_mod6, 
         aes(x = year, y = yvar, color = gndr.c)) +
  geom_smooth(method = "lm", se = FALSE,formula="y~x") +
  geom_point(size=8) +
  geom_point(aes(x=year,y=obs_mean_wt),size=8,shape = 1,alpha=.50)+
  geom_errorbar(aes(ymin=obs_mean_wt_LL, ymax=obs_mean_wt_UL),alpha = .50)+
  scale_color_manual(values = my_colors) +
  scale_y_continuous(limits = c(-1.1, 1.1)) +
  scale_x_continuous(limits = c(2001, 2024),
                     breaks = c(seq(2002, 2020, 2),2023))+
  ylab("Mean-level of value tradition")+
  theme(legend.title=element_blank(),
        legend.text  = element_text(size = 16),
        axis.title.x = element_text(size = 18, face = "bold"),   
        axis.title.y = element_text(size = 18, face = "bold"),   
        axis.text.x  = element_text(size = 14),                  
        axis.text.y  = element_text(size = 14))

p_time_trends

png(filename = 
      "../results/tra/time_trends.png",
    units = "cm",
    width = 21.0,height=29.7*(3/4),res = 300)
p_time_trends
dev.off()


```


### Figures for country-specific time trends

```{r}

# COUNTRY-SPECIFIC TIME x GENDER TRENDS

# obtain country-specific coefficients

mod6_cntry_coefs<-coefficients(mod6)$cntry

# loop through each country and produce a data with all the estimated points at each wave for men and women

countries<-unique(diff_dat$cntry)
pred_list<-list()

for(unique_cntry in countries) {
  unique_cntry_rounds<-
    diff_dat %>%
    filter(cntry == unique_cntry) %>%
    pull(year.c) %>%
    unique()
  
  unique_cntry_coefs<-mod6_cntry_coefs[unique_cntry, ]
  
  unique_cntry_pred<-
    data.frame(cntry=unique_cntry,
               year.c=unique_cntry_rounds,
               gndr.c=rep(x = c(0.5,-0.5),each=length(unique_cntry_rounds)))
  pred_list[[unique_cntry]]<-unique_cntry_pred
}

pred_cntry_dat<-do.call(rbind.data.frame,pred_list)

# model based predictions for each time x country point
pred_cntry_dat$tra.z_mean<-predict(mod6,newdata=pred_cntry_dat)

pred_cntry_dat$year=pred_cntry_dat$year.c+2002

pred_cntry_dat$gender<-
  case_when(
    pred_cntry_dat$gndr.c==0.5~"men",
    pred_cntry_dat$gndr.c==-0.5~"women",
  )

range(pred_cntry_dat$tra.z_mean)

# add observed statistics in addition to model-based estimates
pred_cntry_dat$obs_mean<-NA
pred_cntry_dat$obs_sd<-NA
pred_cntry_dat$obs_n<-NA
pred_cntry_dat$obs_mean_wt<-NA
pred_cntry_dat$obs_sd_wt<-NA
pred_cntry_dat$obs_n_wt<-NA

for(i in 1:nrow(pred_cntry_dat)){
  #i=1
  cntry_i<-pred_cntry_dat[i,"cntry"]
  year_i<-pred_cntry_dat[i,"year"]
  gndr_i<-pred_cntry_dat[i,"gndr.c"]
  
  temp_diff_dat<-diff_dat %>%
    filter(cntry==cntry_i,
           year==year_i,
           gndr.c==gndr_i) %>%
    dplyr::summarize(#n=n(),
      obs_n_wt=sum(pspwght),
      obs_mean_wt=weighted.mean(x=tra.z,w=pspwght),
      obs_sd_wt=sqrt(wtd.var(tra.z,pspwght)),
      obs_mean=mean(tra.z),
      obs_sd=sd(tra.z),
      obs_n=n()) 
  
  pred_cntry_dat[i,"obs_mean"]<-temp_diff_dat[,"obs_mean"]
  pred_cntry_dat[i,"obs_sd"]<-temp_diff_dat[,"obs_sd"]
  pred_cntry_dat[i,"obs_n"]<-temp_diff_dat[,"obs_n"]
  pred_cntry_dat[i,"obs_mean_wt"]<-temp_diff_dat[,"obs_mean_wt"]
  pred_cntry_dat[i,"obs_sd_wt"]<-temp_diff_dat[,"obs_sd_wt"]
  pred_cntry_dat[i,"obs_n_wt"]<-temp_diff_dat[,"obs_n_wt"]
}

# calculate confidence intervals
pred_cntry_dat$obs_mean_se<-pred_cntry_dat$obs_sd/sqrt(pred_cntry_dat$obs_n)
pred_cntry_dat$obs_mean_LL<-pred_cntry_dat$obs_mean+qnorm(.025)*pred_cntry_dat$obs_mean_se
pred_cntry_dat$obs_mean_UL<-pred_cntry_dat$obs_mean+qnorm(.975)*pred_cntry_dat$obs_mean_se

pred_cntry_dat$obs_mean_wt_se<-pred_cntry_dat$obs_sd_wt/sqrt(pred_cntry_dat$obs_n_wt)
pred_cntry_dat$obs_mean_wt_LL<-pred_cntry_dat$obs_mean_wt+qnorm(.025)*pred_cntry_dat$obs_mean_wt_se
pred_cntry_dat$obs_mean_wt_UL<-pred_cntry_dat$obs_mean_wt+qnorm(.975)*pred_cntry_dat$obs_mean_wt_se

# add GEI to figures


index_year_dat<-
  diff_dat_cntry_year %>%
  filter(gndr.c==0.5) %>%
  dplyr::select(cntry,year,gei,gggi,gdi,log_gdp)

pred_cntry_dat<-
  left_join(x=pred_cntry_dat,
            y=index_year_dat,
            by=c("cntry","year"))


# 1) Define scale for GEI
#df_ctry <- pred_cntry_dat[pred_cntry_dat$cntry == ctry, ]

gei_min <- min(pred_cntry_dat$gei, na.rm = TRUE)
gei_max <- max(pred_cntry_dat$gei, na.rm = TRUE)

# Primary axis range
y_min <- -1.1
y_max <-  1.1

# Funktion to transform GEI to primary y-axis
scale_gei_to_y <- function(x) {
  (x - gei_min) / (gei_max - gei_min) * (y_max - y_min) + y_min
}

# Reverse function: primary axis to GEI 
scale_y_to_gei <- function(x) {
  (x - y_min) / (y_max - y_min) * (gei_max - gei_min) + gei_min
}

# save the plots to same pdf-file

#my_colors <- met.brewer("Cassatt2")[c(8, 3)]
my_colors <- met.brewer("Archambault")[c(2,6)]

pdf("../results/tra/country_specific_time_trends.pdf", width = 8, height = 6)

for (ctry in countries) {
  print(
    ggplot(pred_cntry_dat[pred_cntry_dat$cntry == ctry, ],
       aes(x = year, y = tra.z_mean, color = gender)) +
  geom_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  geom_point(size = 8) +
  geom_point(aes(x = year, y = obs_mean_wt), size = 8, shape = 1, alpha = .50) +
  geom_errorbar(aes(ymin = obs_mean_wt_LL, ymax = obs_mean_wt_UL), alpha = .50) +
  geom_line(aes(y = scale_gei_to_y(gei),
                linetype = "GEI"),   
            color = "black",         
            linewidth = 1.2) +
  scale_color_manual(values = my_colors) +
  scale_linetype_manual(          
    name   = NULL,                
    values = c("GEI" = "solid"),
    labels = c("GEI" = "GEI")
  ) +
  geom_flag(aes(country = tolower(ctry))) +
  scale_y_continuous(
    limits = c(y_min, y_max),
    name   = "Mean-level of value tradition",
    sec.axis = sec_axis(~ scale_y_to_gei(.),
                        name = "Gender Equality Index (GEI)")
  ) +
  scale_x_continuous(
    limits = c(2001, 2024),
    breaks = c(seq(2002, 2020, 2), 2023)
  ) +
  ggtitle(paste("Country:", ctry)) +
  theme(legend.title = element_blank())

  )
}

dev.off()


```

### Figure for country-specific trends in panels

```{r}
#ISO<-read.csv2("../data/ISO.csv")

pred_cntry_dat<-left_join(
  x=pred_cntry_dat,
  y=ISO[,c("ISO2","Country_eng_short","CLDR")],
  by=c("cntry"="ISO2")
)

#my_colors <- met.brewer("Cassatt2")[c(8, 3)]
my_colors <- met.brewer("Archambault")[c(2,6)]

facet_plot<-
  ggplot(pred_cntry_dat, 
         aes(x = year, y = tra.z_mean, color = gender)) +
  geom_smooth(method = "lm", se = FALSE, formula="y~x") +
  geom_point(size=3.5) +
  geom_point(aes(x=year,y=obs_mean_wt),size=3.5,shape = 1,alpha=.70)+
  geom_errorbar(aes(ymin=obs_mean_wt_LL, ymax=obs_mean_wt_UL),alpha = .50)+
  geom_line(aes(y = scale_gei_to_y(gei),
                linetype = "GEI"),   
            color = "black",         
            linewidth = 1.2) +
  scale_color_manual(values = my_colors) +
  scale_linetype_manual(          
    name   = NULL,                
    values = c("GEI" = "solid"),
    labels = c("GEI" = "GEI")
  ) +
  #geom_flag(aes(country=tolower(ctry)))+
  scale_y_continuous(
    limits = c(y_min, y_max),
    name   = "Mean-level of value tradition",
    sec.axis = sec_axis(~ scale_y_to_gei(.),
                        name = "Gender Equality Index (GEI)")
  ) +
  scale_x_continuous(limits = c(2001, 2024),
                     breaks = c(seq(2002, 2020, 2),2023)) +
  #ggtitle(paste("Country:", ctry))+
  ylab("Mean-level of value tradition")+
  xlab("Year")+
  theme(legend.title=element_blank(),legend.position = "top",
        axis.text.x = element_text(angle = 45,size = 6,hjust=1))+
  facet_wrap(~CLDR,nrow=6,ncol=6)+
  #facet_wrap(~cntry,nrow=6,ncol=6)+
  geom_flag(aes(country=tolower(cntry)),size=2)

facet_plot

png(filename = 
      "../results/tra/country_time_trend_facets.png",
    units = "cm",
    width = 21.0,height=29.7*(3/4),res = 600)
facet_plot
dev.off()


```

### Exploratory results

Ratio of variances for time main effect and gendered time effect

```{r}
getVC(mod6,round = 10)[3,"vcov"]/
  getVC(mod6,round = 10)[4,"vcov"]

```

Country-specific coefficients for time effect

```{r}
cntry_specific_changes<-
  coefficients(mod6)$cntry %>%
  mutate(change_per_year=year.c,
         gndr_change_per_year=`gndr.c:year.c`) %>%
  mutate(men_change_per_year=change_per_year+0.5*gndr_change_per_year,
         women_change_per_year=change_per_year+(-0.5)*gndr_change_per_year) %>%
  mutate(change_per_21_years=21*change_per_year,
         gndr_change_per_21_year=21*gndr_change_per_year,
         men_change_per_21_years=21*men_change_per_year,
         women_change_per_21_years=21*women_change_per_year) %>%
  dplyr::select(gndr.c,change_per_21_years,gndr_change_per_21_year,men_change_per_21_years,women_change_per_21_years) %>%
  round(.,2)

cntry_specific_changes$cntry<-rownames(cntry_specific_changes)

cntry_specific_changes<-
  left_join(x=cntry_specific_changes,
            y=n_rounds,
            by="cntry")

cntry_specific_changes

# rank by overall change
cntry_specific_changes %>%
  filter(n_unique_essround>4) %>%
  dplyr::select(cntry,change_per_21_years) %>%
  arrange(change_per_21_years)

# rank by gendered change
cntry_specific_changes %>%
  filter(n_unique_essround>4) %>%
  dplyr::select(cntry,gndr_change_per_21_year) %>%
  arrange(gndr_change_per_21_year)

```


#### Country-level predictors: GEI

```{r}


mod6_GEI<-lmer(tra.z~gndr.c+year.c+
             gndr.c:year.c+
               gei.z.cm:gndr.c+gei.z.cm:year.c+gei.z.cm:gndr.c:year.c+
               (gndr.c+year.c+gndr.c:year.c|cntry),
           data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod6_GEI)
getFE(mod6_GEI,round=2,p.round = 5)
getVC(mod6_GEI)

# variance accounted in general slopes
100*(getVC(mod6,round=10)[3,]$vcov-getVC(mod6_GEI,round=10)[3,]$vcov)/
  getVC(mod6,round=10)[3,]$vcov

# variance accounted in gendered slopes
100*(getVC(mod6,round=10)[4,]$vcov-getVC(mod6_GEI,round=10)[4,]$vcov)/
  getVC(mod6,round=10)[4,]$vcov

# GEI-levels  
gei_mean-gei_sd
gei_mean+gei_sd

# Simple slopes for 21 years
change_mod6_GEI<-emmeans(mod6_GEI,specs="year.c",by="gei.z.cm",
                     at=list(gei.z.cm=c(-1,0,1),
                             gndr.c=0,
                             year.c=rev(range(diff_dat$year.c))),
                     disable.pbkrtest=T,
                     lmerTest.limit = 500000,infer=c(T,T))
change_mod6_GEI
pairs(change_mod6_GEI,adjust="none",infer=c(T,T))

# change in gender differences over time

change_in_diff_mod6_GEI<-emmeans(mod6_GEI,specs=c("gndr.c","year.c"),by="gei.z.cm",
                             at=list(gndr.c=c(-0.5,0.5),
                                     gei.z.cm=c(-1,0,1),
                                     year.c=rev(range(diff_dat$year.c))),
                             disable.pbkrtest=T,
                             lmerTest.limit = 500000,infer=c(T,T))
change_in_diff_mod6_GEI
pairs(change_in_diff_mod6_GEI,adjust="none",infer=c(T,T))

# custom contrast to obtain the estimates for the entire span

changes_in_diff <- list(
  diff_ESS11 = c(-1,1,0,0),
  diff_ESS1 = c(0,0,-1,1))

diff_mod6_GEI<-contrast(change_in_diff_mod6_GEI,method = changes_in_diff,adjust="none",infer=c(T,T))

# gender differences at ESS1 and ESS11
diff_mod6_GEI
# Test for whether and how much the differences have changed
pairs(diff_mod6_GEI,infer=c(T,T))


```


#### Country-level predictors: GGGI

```{r}
mod6_GGGI<-lmer(tra.z~gndr.c+year.c+
             gndr.c:year.c+
               gggi.z.cm:gndr.c+gggi.z.cm:year.c+gggi.z.cm:gndr.c:year.c+
               (gndr.c+year.c+gndr.c:year.c|cntry),
           data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod6_GGGI)
getFE(mod6_GGGI,round=2,p.round = 5)
getVC(mod6_GGGI)

# variance accounted in general slopes
100*(getVC(mod6,round=10)[3,]$vcov-getVC(mod6_GGGI,round=10)[3,]$vcov)/
  getVC(mod6,round=10)[3,]$vcov

# variance accounted in gendered slopes
100*(getVC(mod6,round=10)[4,]$vcov-getVC(mod6_GGGI,round=10)[4,]$vcov)/
  getVC(mod6,round=10)[4,]$vcov

# GGGI-levels  
gggi_mean-gggi_sd
gggi_mean+gggi_sd

# Simple slopes for 21 years
change_mod6_GGGI<-emmeans(mod6_GGGI,specs="year.c",by="gggi.z.cm",
                     at=list(gggi.z.cm=c(-1,0,1),
                             gndr.c=0,
                             year.c=rev(range(diff_dat$year.c))),
                     disable.pbkrtest=T,
                     lmerTest.limit = 500000,infer=c(T,T))
change_mod6_GGGI
pairs(change_mod6_GGGI,adjust="none",infer=c(T,T))

# change in gender differences over time

change_in_diff_mod6_GGGI<-emmeans(mod6_GGGI,specs=c("gndr.c","year.c"),by="gggi.z.cm",
                             at=list(gndr.c=c(-0.5,0.5),
                                     gggi.z.cm=c(-1,0,1),
                                     year.c=rev(range(diff_dat$year.c))),
                             disable.pbkrtest=T,
                             lmerTest.limit = 500000,infer=c(T,T))
change_in_diff_mod6_GGGI
pairs(change_in_diff_mod6_GGGI,adjust="none",infer=c(T,T))

# custom contrast to obtain the estimates for the entire span

changes_in_diff <- list(
  diff_ESS11 = c(-1,1,0,0),
  diff_ESS1 = c(0,0,-1,1))

diff_mod6_GGGI<-contrast(change_in_diff_mod6_GGGI,method = changes_in_diff,adjust="none",infer=c(T,T))

# gender differences at ESS1 and ESS11
diff_mod6_GGGI
# Test for whether and how much the differences have changed
pairs(diff_mod6_GGGI,infer=c(T,T))


```

#### Country-level predictors: GDI

```{r}


mod6_GDI<-lmer(tra.z~gndr.c+year.c+
             gndr.c:year.c+
               gdi.z.cm:gndr.c+gdi.z.cm:year.c+gdi.z.cm:gndr.c:year.c+
               (gndr.c+year.c+gndr.c:year.c|cntry),
           data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod6_GDI)
getFE(mod6_GDI,round=2,p.round = 5)
getVC(mod6_GDI)

# variance accounted in general slopes
100*(getVC(mod6,round=10)[3,]$vcov-getVC(mod6_GDI,round=10)[3,]$vcov)/
  getVC(mod6,round=10)[3,]$vcov

# variance accounted in gendered slopes
100*(getVC(mod6,round=10)[4,]$vcov-getVC(mod6_GDI,round=10)[4,]$vcov)/
  getVC(mod6,round=10)[4,]$vcov

# GDI-levels  
gdi_mean-gdi_sd
gdi_mean+gdi_sd

# Simple slopes for 21 years
change_mod6_GDI<-emmeans(mod6_GDI,specs="year.c",by="gdi.z.cm",
                     at=list(gdi.z.cm=c(-1,0,1),
                             gndr.c=0,
                             year.c=rev(range(diff_dat$year.c))),
                     disable.pbkrtest=T,
                     lmerTest.limit = 500000,infer=c(T,T))
change_mod6_GDI
pairs(change_mod6_GDI,adjust="none",infer=c(T,T))

# change in gender differences over time

change_in_diff_mod6_GDI<-emmeans(mod6_GDI,specs=c("gndr.c","year.c"),by="gdi.z.cm",
                             at=list(gndr.c=c(-0.5,0.5),
                                     gdi.z.cm=c(-1,0,1),
                                     year.c=rev(range(diff_dat$year.c))),
                             disable.pbkrtest=T,
                             lmerTest.limit = 500000,infer=c(T,T))
change_in_diff_mod6_GDI
pairs(change_in_diff_mod6_GDI,adjust="none",infer=c(T,T))

# custom contrast to obtain the estimates for the entire span

changes_in_diff <- list(
  diff_ESS11 = c(-1,1,0,0),
  diff_ESS1 = c(0,0,-1,1))

diff_mod6_GDI<-contrast(change_in_diff_mod6_GDI,method = changes_in_diff,adjust="none",infer=c(T,T))

# gender differences at ESS1 and ESS11
diff_mod6_GDI
# Test for whether and how much the differences have changed
pairs(diff_mod6_GDI,infer=c(T,T))


```

#### Country-level predictors: log_GDP

```{r}


mod6_log_GDP<-lmer(tra.z~gndr.c+year.c+
             gndr.c:year.c+
               log_gdp.z.cm:gndr.c+log_gdp.z.cm:year.c+log_gdp.z.cm:gndr.c:year.c+
               (gndr.c+year.c+gndr.c:year.c|cntry),
           data=diff_dat,REML=F,weights = pspwght,
           control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e7)))
summary(mod6_log_GDP)
getFE(mod6_log_GDP,round=2,p.round = 5)
getVC(mod6_log_GDP)

# variance accounted in general slopes
100*(getVC(mod6,round=10)[3,]$vcov-getVC(mod6_log_GDP,round=10)[3,]$vcov)/
  getVC(mod6,round=10)[3,]$vcov

# variance accounted in gendered slopes
100*(getVC(mod6,round=10)[4,]$vcov-getVC(mod6_log_GDP,round=10)[4,]$vcov)/
  getVC(mod6,round=10)[4,]$vcov

# log_GDP-levels  
log_gdp_mean-log_gdp_sd
log_gdp_mean+log_gdp_sd

# Simple slopes for 21 years
change_mod6_log_GDP<-emmeans(mod6_log_GDP,specs="year.c",by="log_gdp.z.cm",
                     at=list(log_gdp.z.cm=c(-1,0,1),
                             gndr.c=0,
                             year.c=rev(range(diff_dat$year.c))),
                     disable.pbkrtest=T,
                     lmerTest.limit = 500000,infer=c(T,T))
change_mod6_log_GDP
pairs(change_mod6_log_GDP,adjust="none",infer=c(T,T))

# change in gender differences over time

change_in_diff_mod6_log_GDP<-emmeans(mod6_log_GDP,specs=c("gndr.c","year.c"),by="log_gdp.z.cm",
                             at=list(gndr.c=c(-0.5,0.5),
                                     log_gdp.z.cm=c(-1,0,1),
                                     year.c=rev(range(diff_dat$year.c))),
                             disable.pbkrtest=T,
                             lmerTest.limit = 500000,infer=c(T,T))
change_in_diff_mod6_log_GDP
pairs(change_in_diff_mod6_log_GDP,adjust="none",infer=c(T,T))

# custom contrast to obtain the estimates for the entire span

changes_in_diff <- list(
  diff_ESS11 = c(-1,1,0,0),
  diff_ESS1 = c(0,0,-1,1))

diff_mod6_log_GDP<-contrast(change_in_diff_mod6_log_GDP,method = changes_in_diff,adjust="none",infer=c(T,T))

# gender differences at ESS1 and ESS11
diff_mod6_log_GDP
# Test for whether and how much the differences have changed
pairs(diff_mod6_log_GDP,infer=c(T,T))


```


# Session information

```{r}
s<-sessionInfo()
print(s,locale=FALSE)
```

