---
title: "Post-hoc power analysis for cross-level interaction versus simple slopes"
output: 
  html_document: 
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To simplify, country-level double-entry data will be used for these purposes. This data matches with Table 1 except no rounding is taking place. 

# Packages

```{r}
library(lme4)
library(rio)
library(multid)
library(emmeans)
library(MASS)
library(matrixcalc)
```

# Data

Get country-level data frame to obtain the parameter estimates of interest for simulation.

```{r}
d<-import("../results/cntry_lvl_data.xlsx")
```

# Parameter estimates

```{r}
# correlation between GGGI and VBMT men
(r_x_y1<-cor(d$gggi.cm,d$`FM M Men`))
# correlation between GGGI and VBMT women
(r_x_y2<-cor(d$gggi.cm,d$`FM M Women`))
# correlation between VBMT men and VBMT women
(r_y1_y2<-cor(d$`FM M Men`,d$`FM M Women`))
# standard deviation of VBMT men
(sd_y1<-sd(d$`FM M Men`))
# standard deviation of VBMT women
(sd_y2<-sd(d$`FM M Women`))
# standard deviation of GGGI
(sd_x<-sd(d$gggi.cm))

```
# Difference score correlation as deconstructed

```{r}
# correlation between GGGI and Difference between men and women in VBMT
(r_x_y1_y2<-cor(d$gggi.cm,d$D))
# calculated from the above parameters
(r_x_y1*sd_y1-r_x_y2*sd_y2)/sqrt(sd_y1^2+sd_y2^2-2*r_y1_y2*sd_y1*sd_y2)
```

These two produce the same number as per Ilmarinen & Lönnqvist (2024)

# Calculate q-estimates

```{r}
# q directly from correlations (not on the same scale because y1 and y2 are standardized separately)
(q<-atanh(r_x_y1)-atanh(r_x_y2))
# q based on slopes that are on the same scale (q_b; y1 and y2 have common standardizer)
# pooled SD
(pooled_sd<-sqrt((sd_y1^2+sd_y2^2)/2))
# SDs harmonized with the pooled SD
(hsd_y1<-sd_y1/pooled_sd)
(hsd_y2<-sd_y2/pooled_sd)
# pooling the harmonized SDs should give SD = 1
sqrt((hsd_y1^2+hsd_y2^2)/2)
# scale the correlations to standardized simple slope estimates
(hr_x_y1<-r_x_y1*hsd_y1)
(hr_x_y2<-r_x_y2*hsd_y2)
# calculate q_b from the standardized simple slopes
(q_b<-atanh(hr_x_y1)-atanh(hr_x_y2))

```
# run a structural path model

```{r}
# needs some renaming
d$Men<-d$`FM M Men`
d$Women<-d$`FM M Women`
d$VBMT<-d$`FM M`

spm<-ddsc_sem(data=d,x="gggi.cm",y1="Men",y2="Women")
round(spm$results,3)

```

Identical estimates to the one's calculated above.

# run a multi-level double entry model

```{r}
# construct a long-format double entry data
long_d<-
  data.frame(
    VBMT=c(d$Men,d$Women),
    gggi.cm=c(d$gggi.cm,d$gggi.cm),
    gndr.c=c(rep(0.5,nrow(d)),rep(-0.5,nrow(d))),
    cntry=c(d$cntry,d$cntry)
    
  )

# standardize gggi.cm
long_d$gggi.z<-(long_d$gggi.cm-mean(d$gggi.cm))/sd_x

# standardize also VBMT with the pooled SD, do not center
long_d$VBMT.z<-(long_d$VBMT)/pooled_sd


mlm<-ddsc_ml(data=long_d,predictor="gggi.z",moderator = "gndr.c",
             moderator_values = c(0.5,-0.5),DV = "VBMT.z",lvl2_unit = "cntry")
round(mlm$results,3)
# check also the model output
fit_mlm<-mlm$model
summary(fit_mlm)
# calculate simple slopes
emtrends(fit_mlm,var="gggi.z",specs="gndr.c",at=list(gndr.c=c(0.5,-0.5)),infer=c(T,T))

```

The deconstructed analysis again provides identical estimates to the above models.



# Power simulation

Run a power simulation from a population which has the same q, q_b, r_x_y1, r_x_y2, sd_y1, sd_y2, sd_x, sd in VBMT between countries, but different r_y1_y2 (correlation between men and women mean-levels). Show that there is more power to detect the cross-level interaction as opposed to the simple slopes of b_11 and b_21 when the correlation is high as it now is in these data (.94)

## Original parameters

```{r}

# covariance matrix
(covmat<-cov(d[,c("gggi.cm","VBMT","Men","Women")]))
# correlation matrix
(cormat<-cor(d[,c("gggi.cm","VBMT","Men","Women")]))
# mean vector
(mu=c(mean(d$gggi.cm),mean(d$VBMT),mean(d$Men),mean(d$Women)))
# sd vector
(sds=c(sd(d$gggi.cm),sd(d$VBMT),sd(d$Men),sd(d$Women)))
# sample size
n=nrow(d)

# save to a list of covariance matrices
cov_mat_list<-list()
cov_mat_list[[1]]<-covmat
```

## Alternative parameters

Change only the covariance between y1 and y2 to lower values: .90, .80, .70, .60, and so on, until .00. Keep the other covariances, variances, etc.

Covariation is a product of correlation and standard deviations: r_y1_y2 x sd_y1 x sd_y2

```{r}
(alt_r_y1_y2<-rev(seq(from=.00,to=.90,by=.10)))

for (i in 1:length(alt_r_y1_y2)){
  alt_covmat<-covmat
  alt_covmat["Men","Women"]<-alt_r_y1_y2[i]*sd_y1*sd_y2
  alt_covmat["Women","Men"]<-alt_r_y1_y2[i]*sd_y1*sd_y2
  cov_mat_list[[1+i]]<-alt_covmat
}

cov_mat_list

# check if each of these are positive definite
for(i in 1:length(alt_r_y1_y2)){
  print(c(i,is.positive.definite(cov_mat_list[[i]])))
}

is.positive.definite(cov_mat_list[[2]])

```

## Power analysis

run 1000 repeats with data generated from each covariance matrix and otherwise equal parameters. Run the double-entry model, and chech the power to detect cross-level interaction and both simple slopes

```{r}


n.sims=10

for(i in 1:n.sims){
  # generate data based on covariance matrix and mean vector
  sim_dat<-data.frame(mvrnorm(n=n,mu = mu,Sigma = cov_mat_list[[i]]))
  # add country (specific countries do not matter here)
  sim_dat$cntry<-d$cntry
  # add general between country variability to means of men and women
  sim_sd_btw<-rnorm(n=n,sd=sd_btw)
  sim_dat$Men<-
  sim_dat$gggi.z<-(sim_dat$gggi.cm-mean(sim_dat$gggi.cm))/sd(sim_dat$gggi.cm)
  sim_pooled_sd<-sqrt((sd(sim_dat$Men)^2+sd(sim_dat$Women)^2)/2)
  
  sim_dat$Men.z<-sim_dat$Men/sim_pooled_sd
  sim_dat$Women.z<-sim_dat$Women/sim_pooled_sd
  
  # construct a long-format double entry data
  sim_long_d<-
  data.frame(
    VBMT.z=c(sim_dat$Men.z,sim_dat$Women.z),
    gggi.z=c(sim_dat$gggi.z,sim_dat$gggi.z),
    gndr.c=c(rep(0.5,nrow(d)),rep(-0.5,nrow(d))),
    cntry=c(sim_dat$cntry,sim_dat$cntry)
    
  )


  sim_mlm<-ddsc_ml(data=sim_long_d,predictor="gggi.z",moderator = "gndr.c",
             moderator_values = c(0.5,-0.5),DV = "VBMT.z",lvl2_unit = "cntry")
  round(sim_mlm$results,3)
  # check also the model output
  fit_sim_mlm<-sim_mlm$model
  summary(fit_sim_mlm)
  # calculate simple slopes
  emtrends(fit_sim_mlm,var="gggi.z",specs="gndr.c",at=list(gndr.c=c(0.5,-0.5)),infer=c(T,T))

  

}

```


