---
title: "Cross-level interaction and simple slope power analysis"
output: 
  html_document: 
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r}
library(ggplot2)
library(ggrepel)
library(MetBrewer)
library(matrixcalc)
library(MASS)
library(lmerTest)
library(emmeans)
library(finalfit)
library(multid)
library(dplyr)
```

# Simulation parameters

Use correlations from .00 to .95 between mean-levels of men (y1) and women (y2).
To use similar effect sizes for cross-level interaction and simple slopes, the association between x and y1 is set at r = .20, association between x and y2 = .00, therefore q is approximately 0.20.

```{r}
# correlations between Y1 and Y2
r_y1_y2<-round(seq(from=.00,to=.95,by=.05),2)
# SD of Y2
sd_y2_vals<-1
# SD of Y1 (kept at constant 1)
sd_y1_vals<-1
# component-specific correlations
r_x_y1=0.20
r_x_y2=0
# compile all combinations of these variable values to the same data frame
simd<-expand.grid(r_y1_y2,sd_y1_vals,sd_y2_vals,r_x_y1,r_x_y2)
# name the columns
names(simd)<-c("r_y1_y2","sd_y1_vals","sd_y2_vals","r_x_y1","r_x_y2")
# calculate the difference score SD
simd$sd_diff_score<-
  sqrt(-2*simd$r_y1_y2*simd$sd_y1_vals*simd$sd_y2_vals+
         simd$sd_y1_vals^2+simd$sd_y2_vals^2)
# calculate the values in the nominator of the difference score correlation equation
# covariation difference
simd$nominator<-(simd$r_x_y1*simd$sd_y1_vals-simd$r_x_y2*simd$sd_y2_vals)
# calculate observed difference score correlation
simd$r_diff_score<-simd$nominator/simd$sd_diff_score
```


## Test if correlation matrices are positive definite

```{r}
posdef<-rep(NA,nrow(simd))

for (i in 1:nrow(simd)){
  testmat<-matrix(c(1,simd$r_x_y1[i],simd$r_x_y2[i],
                    simd$r_x_y1[i],1,simd$r_y1_y2[i],
                    simd$r_x_y2[i],simd$r_y1_y2[i],1), nrow=3, byrow=TRUE )
  posdef[i]<-is.positive.definite(testmat)
}

table(posdef)
```
All 20 matrices are positive definite

## Calculate other summaries for the population matrices

```{r}
# calculate q-estimates (is same for all scenarios)
simd$q=atanh(simd$r_x_y1)-atanh(simd$r_x_y2)
# calculate harmonized SDs
# calculate pooled SD
simd$PSD<-sqrt((simd$sd_y1_vals^2+simd$sd_y2_vals^2)/2)
# calculate harmonized SDs
simd$h_sd_y1_vals<-simd$sd_y1_vals/simd$PSD
simd$h_sd_y2_vals<-simd$sd_y2_vals/simd$PSD
# check that the pooled SDs from these harmonized are all 1
simd$h_PSD<-sqrt((simd$h_sd_y1_vals^2+simd$h_sd_y2_vals^2)/2)
table(round(simd$h_PSD,5)==1)
# calculate harmonized correlations
simd$h_r_x_y1<-simd$r_x_y1*simd$h_sd_y1_vals
simd$h_r_x_y2<-simd$r_x_y2*simd$h_sd_y2_vals
# calculate harmonized q
simd$q_b<-atanh(simd$h_r_x_y1)-atanh(simd$h_r_x_y2)
# print scenarios
round(simd,2)

```


# Simulation

## Function for running the whole thing

```{r}

es_from_mlm<-function(
    s_y1,s_y2,s_x=1, # SDs
    r_y1y2,r_xy1,r_xy2, # correlations
    m_y1=0,m_y2=0,m_x=0, # means
    n=34 # sample size
) {
  
  # mean-vector, all zeros
  mu=c(m_y1,m_y2,m_x)
  # covariance matrix
  cov_mat=matrix(
    c(s_y1^2,r_y1y2*s_y1*s_y2,r_xy1*s_y1*s_x,
      r_y1y2*s_y1*s_y2,s_y2^2,r_xy2*s_y2*s_x,
      r_xy1*s_y1*s_x,r_xy2*s_y2*s_x,s_x^2)
    ,ncol=3,byrow=T)
  # generate data
  d<-data.frame(mvrnorm(n=n,mu=mu,
                        Sigma = cov_mat,empirical = F))
  
  # name variables
  names(d)<-c("y1","y2","x")
  d$id<-1:nrow(d)
  
  # construct a double-entry long-format data.frame
  long_d<-data.frame(
    Y=c(d$y1,d$y2),
    X=c(d$x,d$x),
    W=rep(c(0.5,-0.5),each=nrow(d)),
    ID=c(d$id,d$id)
  )
  
  # fit the cross-level interaction model with the double-entry data
  fit<-lmer(Y~X*W+(1|ID),data=long_d)
  # obtain cross-level interaction
  cl_interaction<-summary(fit)$coefficients["X:W",]
  # obtain simple slopes (unstandardized but does not matter for significance)
  sslopes<-emtrends(fit,var = "X",specs="W",at=list(c(W=c(0.5,-0.5))),infer=c(T,T))
  # save to output
  output<-list(cl_interaction=t(data.frame(cl_interaction)),sslopes=data.frame(sslopes))
  return(output)
}
```


## Run the simulation for each scenario

This is a time consuming part.

```{r message=FALSE, warning=FALSE}
set.seed(11062026)

n.sims <- 2000 # number of repeats per condition
cl_interaction_list <- vector("list", nrow(simd))
sslopes_list        <- vector("list", nrow(simd))
t1<-Sys.time()
for (i in 1:nrow(simd)) {
  cl_interaction_list[[i]] <- vector("list", n.sims)
  sslopes_list[[i]]        <- vector("list", n.sims)
  
  for (j in 1:n.sims) {
    tmp_ES <- es_from_mlm(
      s_y1   = simd[i, "sd_y1_vals"],
      s_y2   = simd[i, "sd_y2_vals"],
      s_x    = 1,
      r_y1y2 = simd[i, "r_y1_y2"],
      r_xy1  = simd[i, "r_x_y1"],
      r_xy2  = simd[i, "r_x_y2"],
      m_y1   = 0,
      m_y2   = 0,
      m_x    = 0,
      n      = 34
    )
    
    cl_interaction_list[[i]][[j]] <- tmp_ES$cl_interaction
    sslopes_list[[i]][[j]]        <- tmp_ES$sslopes
  }
}

t2<-Sys.time()
t2-t1

```

## Compile the results

```{r warning=FALSE}
n.cond <- nrow(simd)

sslopes_df_list<-list()
cl_interaction_df_list<-list()

idx <- 1

for (i in seq_len(n.cond)) {
  for (j in seq_len(n.sims)) {
    
    # Extract stored objects
    sl_obj <- sslopes_list[[i]][[j]]
    cl_obj <- cl_interaction_list[[i]][[j]]
    
    # Coerce to data.frame (1 row each)
    sl_row <- as.data.frame(sl_obj)
    cl_row <- as.data.frame(cl_obj)
    
    # Add condition info (from simd) and simulation index j
    sl_row <- cbind(simd[i, , drop = FALSE],
                    sim = j,
                    sl_row)
    cl_row <- cbind(simd[i, , drop = FALSE],
                    sim = j,
                    cl_row)
    
    sslopes_df_list[[idx]]        <- sl_row
    cl_interaction_df_list[[idx]] <- cl_row
    
    idx <- idx + 1
  }
}

sslopes_df<-bind_rows(sslopes_df_list)
cl_interaction_df<-bind_rows(cl_interaction_df_list)
```

# Results: Power to detect across conditions

```{r}
res_list<-list()

for (i in 1:n.cond){
  temp_cond<-simd[i,"r_y1_y2"]
  
  temp_pow_cl<-sum(cl_interaction_df[cl_interaction_df$r_y1_y2==temp_cond,'Pr(>|t|)']<.10)/n.sims
  temp_pow_y1<-sum(sslopes_df[sslopes_df$r_y1_y2==temp_cond & sslopes_df$W==-0.5,"p.value"]<.10)/n.sims
  temp_pow_y2<-sum(sslopes_df[sslopes_df$r_y1_y2==temp_cond & sslopes_df$W==0.5,"p.value"]<.10)/n.sims
  res_list[[i]]<-c(cor_btw_y1_and_y2=temp_cond,
                   Power_cross_level_interaction=temp_pow_cl,
                   False_positive_rate_y1_slope=temp_pow_y1,
                   Power_y2_slope=temp_pow_y2)


}

res_df<-data.frame(do.call(rbind,res_list))
res_df

```

## Figure

```{r}
# stack the power estimates

plot_dat<-
  data.frame(
    cor_btw_y1_and_y2=rep(res_df$cor_btw_y1_and_y2,2),
    power_est=c(res_df$Power_cross_level_interaction,
                res_df$Power_y2_slope),
    est_type=c(rep(c("Cross-level interaction: q = 0.20","Simple slope: r = 0.20"),each=nrow(res_df)))
  )

# add correlations for VBMT and ten values
cor_points<-
  data.frame(
    value=c("VBMT","hedonism","self-direction","security","tradition","conformity","power",
            "stimulation","benevolence","achievement","universalism"),
    cor_btw_y1_and_y2=c(.94,.96,.92,.98,.96,.95,.99,
                  .95,.96,.98,.95)
  )

cor_points

plot <- ggplot(
  data = plot_dat,
  aes(
    y = power_est,
    x = cor_btw_y1_and_y2,
    color = est_type,
    fill = est_type
  )
) +
  geom_line(linewidth = 2) +

  geom_vline(
    data = cor_points,
    aes(xintercept = cor_btw_y1_and_y2),
    colour = "grey60",
    linetype = "solid",
    inherit.aes = FALSE
  ) +

  # bracket line
  annotate(
    "segment",
    x = .92, xend = .99,
    y = 1.2, yend = 1.2,
    linewidth = 0.8
  ) +

  # bracket ends
  annotate(
    "segment",
    x = .92, xend = .92,
    y = 1.15, yend = 1.2,
    linewidth = 0.8
  ) +
  annotate(
    "segment",
    x = .99, xend = .99,
    y = 1.15, yend = 1.2,
    linewidth = 0.8
  ) +

  # annotation text
  annotate(
    "text",
    x = .955,
    y = 1.28,
    label = "VBMT and all ten value correlations\nfall within this range (.92-.99)",
    size = 4
  ) +

  ylab("Statistical power to detect with alpha = .10 and n = 34") +
  xlab("Correlation between mean-levels of men and women") +

  scale_x_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = .10)
  ) +
  scale_y_continuous(
    limits = c(0, 1.3),
    breaks = seq(0, 1, by = .10)
  ) +

  coord_cartesian(
    ylim = c(0, 1.08),
    clip = "off"
  ) +

  theme(
    legend.title = element_blank(),
    plot.margin = margin(t = 60)
  )

plot


png(filename = 
      "../results/Power_analysis_comparison.png",
    units = "cm",
    width = 21.0,height=29.7*(2/4),res = 300)
plot
dev.off()
```

# Session information

```{r}
s<-sessionInfo()
print(s,locale=FALSE)
```
