---
title: "Data preparations for European Social Survey"
output: 
  html_document: 
    toc: true
    keep_md: true
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r message=FALSE}
library(rio)
library(dplyr)
```

# Read Data

This reads directly from the zip

```{r}

zip_path <- "../data/ESS-data_subset.zip"
tmp_dir  <- tempdir()

unzip(zip_path, files = "Datafile-subset.sav", exdir = tmp_dir, overwrite = TRUE)

d <- import(file.path(tmp_dir, "Datafile-subset.sav"))

table(d$essround,useNA="always")
table(d$cntry,useNA="always")
```

# Calculate the number of waves per each country

```{r}
gd<-data.frame(cntry=unique(d$cntry))
gd$waves<-NA

for (i in 1:nrow(gd)){
  gd[i,"waves"]<-length(unique(d[d$cntry==unique(d$cntry)[i],
                                 "essround"]))
}
gd  

# attach number of rounds variable to data

d<-left_join(
  x=d,
  y=gd,
  by="cntry")

table(d$waves,useNA="always")
```

# Gender

```{r}

# code binary gender variable
attributes(d$gndr)
table(d$gndr,useNA="always")
d$gndr.bin<-ifelse(d$gndr==2,0,d$gndr)
table(d$gndr.bin,useNA="always")

# code effect coded gender variable
d$gndr.c<-d$gndr.bin-0.5
table(d$gndr.c,useNA="always")
```

# Childlessness

```{r}
# first based only on the single variable
# this variable has a large number of missing values
# because it is a follow up question
# but the prior question is only present in ESS1-8
attributes(d$chldhhe)
table(d$chldhhe,useNA="always")

table(d$essround,d$chldhhe,useNA="always")
prop.table(table(d$essround,d$chldhhe,useNA="always"),1)

d$childless1<-
  case_when(d$chldhhe==2~1,
            d$chldhhe==1~0,
            TRUE~NA_integer_)
table(d$childless1,useNA="always")
prop.table(table(d$childless1,useNA="always"))
prop.table(table(d$childless1))

# then based on the two variables
# this latter includes those who live with a kid currently

attributes(d$chldhm)
table(d$chldhhe,d$chldhm,useNA="always")

d$childless2<-
  case_when(d$chldhhe==2 & d$chldhm==2~1, # childless if ever and currently childless
            d$chldhhe==1 | d$chldhm==1 ~0, # non-childless if ever OR currently has kids in the HH
            TRUE~NA_integer_)
table(d$childless2,useNA="always")
prop.table(table(d$childless2,useNA="always"))
prop.table(table(d$childless2))

table(d$essround,d$childless2,useNA="always")

# check the responses in rshipa
attributes(d$rship10)
attributes(d$rshipa20)

hh_rshipa=paste0("rshipa",2:24)
hh_rship=paste0("rship",2:9)

d$hh_child1<-apply(d[,hh_rship], 1, function(row) as.integer(any(row == 2)))
table(d$hh_child1,useNA="always")

d$hh_child2<-apply(d[,hh_rshipa], 1, function(row) as.integer(any(row == 2)))
table(d$hh_child2,useNA="always")

d$hh_child<-ifelse(d$hh_child1==1 | d$hh_child2==1,1,0)
table(d$hh_child,useNA="always")
# code NA as 0
d$hh_child[is.na(d$hh_child)]<-0
table(d$hh_child,useNA="always")

table(d$chldhm,d$hh_child,useNA="always")

# try to apply this same system for all waves
# if there is a child in the HH, then the person in not childless
# if there is not a child in the HH, then the second question is used
# to define if the person is childless


d$childless3<-
  case_when(d$hh_child==1 | d$chldhhe==1~0, # non-childless if kid at home OR ever had
            d$hh_child==0 & d$chldhhe==2~1, # childless if no kid at household AND never was
            TRUE~NA_integer_)
table(d$childless3,useNA="always")
prop.table(table(d$childless3,useNA="always"))
prop.table(table(d$childless3))
```

# essround

```{r}
# center around midpoint (first round + last round)/2 = (1+11)/2 = 6
d$essround.c<-d$essround-6
table(d$essround.c)
```

# Age

```{r}
attributes(d$agea)

d$age.c<-scale(d$agea,center = T,scale=T)

# belong to the focal age range of childlessness examination

d$age_included<-
  case_when(d$agea>17 & d$agea<46~1,
            TRUE~0)
table(d$age_included,useNA="always")

table(d$essround,
      d$age_included,useNA="always")
```

# Education

```{r}
attributes(d$eduyrs)
table(d$eduyrs,useNA="always")

d$eduyrs.c<-scale(d$eduyrs,center = T,scale=T)
```

# Religiosity

```{r}
attributes(d$rlgdgr)
table(d$rlgdgr,useNA="always")
d$rlgdgr.c<-scale(d$rlgdgr,center=T,scale=T)

# country mean-level religiosity

reld<-d %>%
  group_by(cntry) %>%
  summarise(rlgdgr.cm=mean(rlgdgr.c,na.rm=T))
reld

d<-
  left_join(
    x=d,
    y=reld,
    by="cntry")
```

# Country x time datasets

```{r}
d$cntry_time<-paste0(d$cntry,"_",d$essround)
table(d$cntry_time,useNA="always")
```

# Same-gender partnership

```{r}
# must be done separately for ESS1

# use a function that checks if there are 
# same-sex partners living in the household

has_samegndr_partner<-function(data,hh_gndr,hh_rshipa){
  
  gndr_self<-data[,"gndr"]
  test_mat<-matrix(NA,nrow=nrow(data),ncol=length(hh_gndr))
  
  for (i in 1:length(hh_gndr)){
    test_mat[,i]<-
      (gndr_self == data[,hh_gndr[i]] & data[,hh_rshipa[i]]==1)
  }
  
  result<-rowSums(test_mat,na.rm=T)
  return(result)
  
}


hh_gndr=paste0("gndr",2:24)
hh_rshipa=paste0("rshipa",2:24)

d$same_gndr210<-has_samegndr_partner(data=d,
                        hh_gndr=hh_gndr,
                        hh_rshipa=hh_rshipa)
table(d$same_gndr210,useNA="always")

d$same_gndr1<-has_samegndr_partner(
  data=d,
  hh_gndr=paste0("gndr",2:9),
  hh_rshipa=paste0("rship",2:9)
)

table(d$same_gndr1,useNA="always")

# combine

d$same_gndr_partner<-
  case_when(d$same_gndr1>0 | d$same_gndr210>0~1,
            TRUE~0)
table(d$same_gndr_partner,useNA="always")
```

# Values

For rounds 1-10 the variable names are the same. For round 11, the variables have an additional "a" at the end. First merge them to same variables.

```{r}
table(d$essround,is.na(d$impdiff)) 
table(d$essround,is.na(d$impdiffa))

table(!is.na(d$impdiff),!is.na(d$impdiffa))

value_items <- c(
  "impdiff",
  "impenv",
  "impfree",
  "impfun",
  "imprich",
  "impsafe",
  "imptrad",
  "ipadvnt",
  "ipbhprp",
  "ipcrtiv",
  "ipeqopt",
  "ipfrule",
  "ipgdtim",
  "iphlppl",
  "iplylfr",
  "ipmodst",
  "iprspot",
  "ipshabt",
  "ipstrgv",
  "ipsuces",
  "ipudrst"
)
```

```{r}

for (x in value_items) {
  var_a   <- paste0(x, "a")
  var_out <- paste0(x, "_comb")
  d[[var_out]] <- ifelse(!is.na(d[[var_a]]), d[[var_a]], d[[x]])
}

table(d$essround,is.na(d$impdiff_comb))

```



```{r}
## reverse code all value items

d$impdiff.R<-7-d$impdiff_comb
d$impenv.R<-7-d$impenv_comb
d$impfree.R<-7-d$impfree_comb
d$impfun.R<-7-d$impfun_comb
d$imprich.R<-7-d$imprich_comb
d$impsafe.R<-7-d$impsafe_comb
d$imptrad.R<-7-d$imptrad_comb
d$ipadvnt.R<-7-d$ipadvnt_comb
d$ipbhprp.R<-7-d$ipbhprp_comb
d$ipcrtiv.R<-7-d$ipcrtiv_comb
d$ipeqopt.R<-7-d$ipeqopt_comb
d$ipfrule.R<-7-d$ipfrule_comb
d$ipgdtim.R<-7-d$ipgdtim_comb
d$iphlppl.R<-7-d$iphlppl_comb
d$iplylfr.R<-7-d$iplylfr_comb
d$ipmodst.R<-7-d$ipmodst_comb
d$iprspot.R<-7-d$iprspot_comb
d$ipshabt.R<-7-d$ipshabt_comb
d$ipstrgv.R<-7-d$ipstrgv_comb
d$ipsuces.R<-7-d$ipsuces_comb
d$ipudrst.R<-7-d$ipudrst_comb
```

Calculate the ten value aggregates from the combined and reverse scored items

```{r}
# conformity
attributes(d$ipfrule)
attributes(d$ipbhprp)
d$con<-rowMeans(d[,c("ipfrule.R","ipbhprp.R")],na.rm=T)

# tradition
attributes(d$ipmodst)
attributes(d$imptrad)
d$tra<-rowMeans(d[,c("ipmodst.R","imptrad.R")],na.rm=T)

# benevolence

attributes(d$iphlppl)
attributes(d$iplylfr)
d$ben<-rowMeans(d[,c("iphlppl.R","iplylfr.R")],na.rm=T)

# universalism

attributes(d$ipeqopt)
attributes(d$ipudrst)
attributes(d$impenv)
d$uni<-rowMeans(d[,c("ipeqopt.R","ipudrst.R","impenv.R")],na.rm=T)

# self-diretion

attributes(d$ipcrtiv)
attributes(d$impfree)
d$sdi<-rowMeans(d[,c("ipcrtiv.R","impfree.R")],na.rm=T)

# stimulation

attributes(d$impdiff)
attributes(d$ipadvnt)
d$sti<-rowMeans(d[,c("impdiff.R","ipadvnt.R")],na.rm=T)

# hedonism

attributes(d$ipgdtim)
attributes(d$impfun)
d$hed<-rowMeans(d[,c("ipgdtim.R","impfun.R")],na.rm=T)

# achievement

attributes(d$ipshabt)
attributes(d$ipsuces)
d$ach<-rowMeans(d[,c("ipshabt.R","ipsuces.R")],na.rm=T)

# power

attributes(d$imprich)
attributes(d$iprspot)
d$pow<-rowMeans(d[,c("imprich.R","iprspot.R")],na.rm=T)

# security

attributes(d$impsafe)
attributes(d$ipstrgv)
d$sec<-rowMeans(d[,c("impsafe.R","ipstrgv.R")],na.rm=T)
```

## Within-country standardized values

```{r}

# center individual scores around country mean-levels on values
cntry_values<-d %>%
  group_by(cntry) %>%
  summarise(con.cm=mean(con,na.rm=T),
            tra.cm=mean(tra,na.rm=T),
            ben.cm=mean(ben,na.rm=T),
            uni.cm=mean(uni,na.rm=T),
            sdi.cm=mean(sdi,na.rm=T),
            sti.cm=mean(sti,na.rm=T),
            hed.cm=mean(hed,na.rm=T),
            ach.cm=mean(ach,na.rm=T),
            pow.cm=mean(pow,na.rm=T),
            sec.cm=mean(sec,na.rm=T),
            con.csd=sd(con,na.rm=T),
            tra.csd=sd(tra,na.rm=T),
            ben.csd=sd(ben,na.rm=T),
            uni.csd=sd(uni,na.rm=T),
            sdi.csd=sd(sdi,na.rm=T),
            sti.csd=sd(sti,na.rm=T),
            hed.csd=sd(hed,na.rm=T),
            ach.csd=sd(ach,na.rm=T),
            pow.csd=sd(pow,na.rm=T),
            sec.csd=sd(sec,na.rm=T))

# add the means and sds to original data frame
d<-left_join(
  x=d,
  y=cntry_values,
  by="cntry"
)  


d$con.cz<-(d$con-d$con.cm)/d$con.csd
d$tra.cz<-(d$tra-d$tra.cm)/d$tra.csd
d$ben.cz<-(d$ben-d$ben.cm)/d$ben.csd
d$uni.cz<-(d$uni-d$uni.cm)/d$uni.csd
d$sdi.cz<-(d$sdi-d$sdi.cm)/d$sdi.csd
d$sti.cz<-(d$sti-d$sti.cm)/d$sti.csd
d$hed.cz<-(d$hed-d$hed.cm)/d$hed.csd
d$ach.cz<-(d$ach-d$ach.cm)/d$ach.csd
d$pow.cz<-(d$pow-d$pow.cm)/d$pow.csd
d$sec.cz<-(d$sec-d$sec.cm)/d$sec.csd

```

## Within-country standardized value items

```{r}

cntry_items<-d %>%
  group_by(cntry) %>%
  summarise(
    impdiff.R.cm=mean(impdiff.R,na.rm=T),
    impenv.R.cm=mean(impenv.R,na.rm=T),
    impfree.R.cm=mean(impfree.R,na.rm=T),
    impfun.R.cm=mean(impfun.R,na.rm=T),
    imprich.R.cm=mean(imprich.R,na.rm=T),
    impsafe.R.cm=mean(impsafe.R,na.rm=T),
    imptrad.R.cm=mean(imptrad.R,na.rm=T),
    ipadvnt.R.cm=mean(ipadvnt.R,na.rm=T),
    ipbhprp.R.cm=mean(ipbhprp.R,na.rm=T),
    ipcrtiv.R.cm=mean(ipcrtiv.R,na.rm=T),
    ipeqopt.R.cm=mean(ipeqopt.R,na.rm=T),
    ipfrule.R.cm=mean(ipfrule.R,na.rm=T),
    ipgdtim.R.cm=mean(ipgdtim.R,na.rm=T),
    iphlppl.R.cm=mean(iphlppl.R,na.rm=T),
    iplylfr.R.cm=mean(iplylfr.R,na.rm=T),
    ipmodst.R.cm=mean(ipmodst.R,na.rm=T),
    iprspot.R.cm=mean(iprspot.R,na.rm=T),
    ipshabt.R.cm=mean(ipshabt.R,na.rm=T),
    ipstrgv.R.cm=mean(ipstrgv.R,na.rm=T),
    ipsuces.R.cm=mean(ipsuces.R,na.rm=T),
    ipudrst.R.cm=mean(ipudrst.R,na.rm=T),
    impdiff.R.csd=sd(impdiff.R,na.rm=T),
    impenv.R.csd=sd(impenv.R,na.rm=T),
    impfree.R.csd=sd(impfree.R,na.rm=T),
    impfun.R.csd=sd(impfun.R,na.rm=T),
    imprich.R.csd=sd(imprich.R,na.rm=T),
    impsafe.R.csd=sd(impsafe.R,na.rm=T),
    imptrad.R.csd=sd(imptrad.R,na.rm=T),
    ipadvnt.R.csd=sd(ipadvnt.R,na.rm=T),
    ipbhprp.R.csd=sd(ipbhprp.R,na.rm=T),
    ipcrtiv.R.csd=sd(ipcrtiv.R,na.rm=T),
    ipeqopt.R.csd=sd(ipeqopt.R,na.rm=T),
    ipfrule.R.csd=sd(ipfrule.R,na.rm=T),
    ipgdtim.R.csd=sd(ipgdtim.R,na.rm=T),
    iphlppl.R.csd=sd(iphlppl.R,na.rm=T),
    iplylfr.R.csd=sd(iplylfr.R,na.rm=T),
    ipmodst.R.csd=sd(ipmodst.R,na.rm=T),
    iprspot.R.csd=sd(iprspot.R,na.rm=T),
    ipshabt.R.csd=sd(ipshabt.R,na.rm=T),
    ipstrgv.R.csd=sd(ipstrgv.R,na.rm=T),
    ipsuces.R.csd=sd(ipsuces.R,na.rm=T),
    ipudrst.R.csd=sd(ipudrst.R,na.rm=T))

# add the means and sds to original data frame
d<-left_join(
  x=d,
  y=cntry_items,
  by="cntry"
)  

d$impdiff.R.cz<-(d$impdiff.R-d$impdiff.R.cm)/d$impdiff.R.csd
d$impenv.R.cz<-(d$impenv.R-d$impenv.R.cm)/d$impenv.R.csd
d$impfree.R.cz<-(d$impfree.R-d$impfree.R.cm)/d$impfree.R.csd
d$impfun.R.cz<-(d$impfun.R-d$impfun.R.cm)/d$impfun.R.csd
d$imprich.R.cz<-(d$imprich.R-d$imprich.R.cm)/d$imprich.R.csd
d$impsafe.R.cz<-(d$impsafe.R-d$impsafe.R.cm)/d$impsafe.R.csd
d$imptrad.R.cz<-(d$imptrad.R-d$imptrad.R.cm)/d$imptrad.R.csd
d$ipadvnt.R.cz<-(d$ipadvnt.R-d$ipadvnt.R.cm)/d$ipadvnt.R.csd
d$ipbhprp.R.cz<-(d$ipbhprp.R-d$ipbhprp.R.cm)/d$ipbhprp.R.csd
d$ipcrtiv.R.cz<-(d$ipcrtiv.R-d$ipcrtiv.R.cm)/d$ipcrtiv.R.csd
d$ipeqopt.R.cz<-(d$ipeqopt.R-d$ipeqopt.R.cm)/d$ipeqopt.R.csd
d$ipfrule.R.cz<-(d$ipfrule.R-d$ipfrule.R.cm)/d$ipfrule.R.csd
d$ipgdtim.R.cz<-(d$ipgdtim.R-d$ipgdtim.R.cm)/d$ipgdtim.R.csd
d$iphlppl.R.cz<-(d$iphlppl.R-d$iphlppl.R.cm)/d$iphlppl.R.csd
d$iplylfr.R.cz<-(d$iplylfr.R-d$iplylfr.R.cm)/d$iplylfr.R.csd
d$ipmodst.R.cz<-(d$ipmodst.R-d$ipmodst.R.cm)/d$ipmodst.R.csd
d$iprspot.R.cz<-(d$iprspot.R-d$iprspot.R.cm)/d$iprspot.R.csd
d$ipshabt.R.cz<-(d$ipshabt.R-d$ipshabt.R.cm)/d$ipshabt.R.csd
d$ipstrgv.R.cz<-(d$ipstrgv.R-d$ipstrgv.R.cm)/d$ipstrgv.R.csd
d$ipsuces.R.cz<-(d$ipsuces.R-d$ipsuces.R.cm)/d$ipsuces.R.csd
d$ipudrst.R.cz<-(d$ipudrst.R-d$ipudrst.R.cm)/d$ipudrst.R.csd

```

# Marriage

```{r}
attributes(d$marital)
table(d$marital,useNA="always")

attributes(d$maritala)
table(d$maritala,useNA="always")

attributes(d$maritalb)
table(d$maritalb,useNA="always")

d$married<-case_when(d$marital==1 |
                    d$maritala==1 |
                    d$maritalb==1 ~1,
                    TRUE~0)
table(d$married,useNA="always")

# mid-point centered marriage
d$married.c<-d$married-0.5
```

# Domicile

```{r}
attributes(d$domicil)
table(d$domicil,useNA="always")
d$rural=case_when(
  d$domicil==4 | d$domicil==5~1,
  d$domicil<4~0,
  TRUE~NA_integer_
)
table(d$rural,useNA="always")

d$rural.c<-d$rural-0.5
table(d$rural.c,useNA="always")

```

# Variable selection

```{r}

fdat<-d %>%
  dplyr::select(
    essround,essround.c,idno,cntry,waves,pspwght,cntry_time,
    rlgdgr,rlgdgr.c,rlgdgr.cm,
    gndr,gndr.bin,gndr.c,
    agea,age.c,age_included,
    chldhhe,chldhm,childless1,childless2,childless3,
    eduyrs,eduyrs.c,
    married,married.c,
    rural,rural.c,
    same_gndr_partner,
    con,tra,ben,uni,sdi,sti,hed,ach,pow,sec
  )
```

# Data export

```{r}
save(fdat,file="../data/fdat.rdata")
```

# Session information

```{r}
s<-sessionInfo()
print(s,locale=F)
```